%! program = pdflatex

\documentclass[12pt,letter]{memoir} % for a long document
%\documentclass[12pt,a4paper,article]{memoir} % for a short document

%\usepackage[T1]{fontenc}
\usepackage{bookman}

% See the ``Memoir customise'' template for some common customisations
% Don't forget to read the Memoir manual: memman.pdf

\title{Code of the Illuminati\\
	Part III of\\
	Memoirs Illustrating the History of Jacobinism}
\author{Abbe Barreul\\
	translated by\\
	Robert Edward Clifford}
\date{1798} % Delete this line to display the current date

%%% BEGIN DOCUMENT
\begin{document}

\maketitle
\tableofcontents* % the asterisk means that the contents itself isn't put into the ToC

\chapter{Preliminary Observations}
\textit{On the Illuminees\footnote{The Translator thinks it proper to inform the Reader, that, considering how much the abuse of terms, such as of Philosophy, Reason, \&c. \&c. has contributed to diffuse the new-fangled doctrines, he has adopted in the present volume (which may be said to be the first methodical work published on the subject of which it treats) the words Illumine, Illuminize, and Illuminization, though Illuminate and Illumination might perhaps be more correct expressions. Every reader will feel, that the illumination of the world, and to illuminate mankind, are objects worthy of the true philosopher. But may the man be ever accurst who shall attempt to illuminize his countrymen, or aim at the illuminization of the world! T.} and on the different Works whereon these Memoirs are grounded.}

The third conspiracy, which I am now about to investigate, is that of the Atheistical Illuminees, which at my outset\footnote{Vol. I. page xxii.} I denominated the conspiracy of the Sophisters of Impiety and Anarchy against every religion natural or revealed; not only against kings, but against every government, against all civil society, even against all property whatsoever.

The name of Illuminee which this Sect (the most disastrous in its principles, the most extensive in its views, the most atrociously cunning in its means) has chosen, is of ancient standing in the annals of disorganizing Sophistry. It was the name which Manes and his disciples first affected, gloriantur Manichï¿½i se de cï¿½lo illuminatos\footnote{Gaultier, Verbo Manichï¿½i, Sect. 3.}.  The first Rosicrucians also, who appeared in Germany, called themselves Illuminees. And later, in our time, the Martinists (with many other sects) have pretended to Illuminism. As an outline for history I distinguish them by their plots and tenets, and will reduce them into two classes, the Atheistical and the Theosophical Illuminees. These latter more particularly comprehend the Martinists, whom I have already mentioned in my second volume, and the Swedenbourgians, whom I shall mention in their proper place, where also I shall give what information I have been able to collect relating to them. The Atheistical Illuminees are the objects of the present volume, and it is their conspiracy that I mean to disclose.

The very numerous letters, books, and manuscripts, which I have received since the publication of my proposals, has rendered it impossible for me to comprise the proposed investigation in one volume. The baleful projects of the Sect and the laws for their execution are so strangely combined, that I thought it necessary to begin by making my reader perfectly acquainted with its code; that is to say, with the regular progression of its degrees, mysteries, and government.

This alone requiring an entire volume, I am reduced to the necessity of giving a fourth, in which I shall develope the history of Illuminism, and make

p. 394

an application of the triple conspiracy to the French Revolution. I have more particularly applied myself to the investigation of the legislative part of this conspiring Sect, as no work has yet been published in which the whole of their code is to be found. Detached parts only were to be met with scattered throughout the papers which had been seized by the public authority. These I have collected and digested; thus enabling the reader more easily to judge what has been and what must have been the result of such laws. In such an undertaking, I feel myself bound to lay before the public an account of the documents on which I ground my proofs. The following then is a list of the principal works, with a few observations on each, that the reader may form his own judgment as to their authenticity.

I. The first is a collection entitled "Some of the Original Writings of the Sect of Illuminees, which were discovered on the 11th and 12th of October, 1786, at Landshut, on a search made in the House of the Sieur Zwack, heretofore Counsellor of the Regency; and printed by Order of His Highness the Elector.ï¿½Munich, by Ant. Franz, Printer to the Court."\footnote{Einige original schriften des Illuminaten Ordens, welche bey dem gewesenen regierungsrath Zwack, durch vorgennommene haus visitation zu Landshut den 11 and 12 Octob. 1786, vorgefunden worden. Auf hï¿½chsten befehl seiner churfï¿½rstlichen Durchleucht zum druck befï¿½rdert. Mï¿½nchen. Gedruckt bey Ant Franz churfl; hof-buchdrucker.}

II. The second is a supplement to the Original Writings, chiefly containing those which were found on a search made at the castle of Sandersdorf, a famous haunt of the Illuminees, by order of His Highness the Elector. Munich, 1787.\footnote{Nachrichten von weitern Original schriften, \&c. \&c.}

These two volumes contain irrefragable proofs of the most detestable conspiracy. They disclose the principles, the object, and the means of the Sect; the essential parts of their code, the diligent correspondence of the adepts, particularly that of their chief, and a statement of their progress and future hopes. The editors indeed have carried their attention so far, as to mention by whose hand the principal documents or letters were written. At the beginning of the first volume, and on the frontispiece of the second, is seen the following remarkable advertisement by order of the Elector:ï¿½"Those who may harbour any doubt as to the authenticity of this collection, have only to apply to the office where the secret archives are kept at Munich, and where orders are left to show the originals."\footnote{Wer an der aechtheit dieser versammlung einen zweifel trï¿½gt, mag sich nur bey den hiesigen geheimen archiv melden, allwo man ihm die urschrifften selbst vorzulegen befehligen ist. Mï¿½nchen 26 Mï¿½rz 1787.}

I entreat that my readers will recollect this advertisement whenever they shall see the Original Writings cited.

III. "The True Illuminee, or the real and perfect Ritual of the Illuminee; of the Illuminee; comprehending the Preparation, the Noviciate, the Minerval Degree, that of the Minor and Major Illuminee, all without addition or omission."ï¿½With respect to the authenticity of this work, we need only quote the testimony of the Baron Knigge, surnamed Philo, the most famous of the Illuminees after the Founder of the Sect; and who was actually the chief compiler of its Code, as he tells us himself: "All these degrees (says he), such as I composed them, have been printed this year at Edesse (Frankfort on the Mein) under the title of the True Illuminee. I am ignorant of the author; but they appear exactly as they flowed from my pen; that is to say, as I compiled them."\footnote{Philo's Endliche erklï¿½rung, \&c. Page 96.}  This certainly is an authenticated document on the Sect, and recognized by the compiler himself.

p. 395

IV. I now proceed to a work which was published by this same Philo,ï¿½under the title of "Last Observations, or Last Words of Philo, and Answers to divers Questions on my connections with the Illuminees." In this work Philo-Knigge gives us an account of himself and of his Illuminism, of his agreements with the chiefs of the Sect, and of his labours for it. His vanity, however, makes this narrative fulsome. The reader will observe in his writings one of those pretended Philosophers who treat all religious objects with that contempt which they themselves deserve. This is of no consequence; he attempts to justify his own conduct; his avowals may therefore be received in testimony against the Sect.

V. "The last Works of Spartacus and Philo," Die neusten Arbeiten des Spartacus und Philo. Except the Original Writings, this is the most intelligent and important work that has been published on the Illuminees. It contains the two degrees of the greatest consideration both on account of the mysteries revealed in them by the Sect, and of the laws laid down for the adepts.ï¿½Not a shadow of doubt can be maintained as to the authenticity of this work. These degrees and laws are published with a certificate of Philo attesting their conformity with the original, and under the seal of the Order. This certificate was scarcely necessary. Whoever can read must easily perceive that these degrees and these laws are no other than a compilation, and often (in the most essential parts) but a copy of the discourses, precepts, and principles, contained in the Original Writings. The publisher is a man who has passed through all the degrees of Illuminism. More dexterous than Philo, he makes himself master of his secret, and of that of the whole Sect. The better to unmask Illuminism, he becomes an Illuminee; and he has so well succeeded, that no member of the Order was better acquainted with it than himself.

VI. The same writer has published A Critical History of the Degrees of Illuminism, a valuable work, in which every thing is proved from the very letters of the grand adepts.

VII. The Directing Illuminee, or the Scotch Knight. This may be said to be the counterpart of the Last Works of Philo and Spartacus. It is a description of the most important intermediary degree of Illuminism. The Editor does not indeed publish it under the signet of the Order; but when the reader has compared it with the Original Writings, and even with the criticism on it by the chief, who was not much pleased with the compiler, he will soon decide that the grand seal of the Order is not necessary to authenticate it.

VIII. Remarkable Depositions respecting the Illuminees. These are three juridical depositions on oath, and signed 1st by Mr. Cosandy, Canon and Professor at Munich; 2dly by Mr. Renner, Priest and Professor of the same Academy; 3dly by Mr. Utzschneider, Counsellor of the Electoral Chamber; 4thly by Mr. George Grï¿½mberg, a member of the Academy of Sciences, and Professor of Mathematics. As every thing is juridical in these depositions, it would be useless for me to insist on the weight they must carry with them. These were four pupils, who did not wait to be initiated in the grand mysteries of the Sect to form their judgement on, and to quit the Sect. They

p. 396

were cited at a tribunal to declare all they knew, and they answered with moderation and truth. Their depositions will find a place in the historical part of this work.

IX. The Apologies published by some of the leaders of the Sect are also to be classed among the incontrovertible evidence which we have acquired. These gentlemen will not be expected to have aggravated their own wickedness.

X. The list would be endless were I to subjoin all the works that have been written against the Sect. But I must distinguish in this place the works of Mr. Hoffman, Professor at the University of Vienna. I am but little acquainted with those of Doctor Zimmerman, though I have been informed by letter, that he furnished many valuable articles in a journal published at Vienna, and chiefly directed against the Sect. I often find Mr. Starkï¿½s name mentioned as a strenuous opponent of the Sect. I have seen no publication with his name to it, except an Apology in Answer to the Calumnies of the Sect, which it continues to repeat, notwithstanding the victorious manner in which he has answered them.

Among the anonymous writings I find an excellent work entitled the Ultimate fate of the Free-masons (Endliches schicksal des Frey-maurer Ordens). It is a discourse pronounced at the breaking-up of a Freemason's Lodge. The writer of this discourse gives an excellent statement of the reasons why the Lodges should suspend their labours since Illuminism had intruded itself into Masonry.ï¿½I believe he would have pronounced this discourse much sooner, had he known that all Lodges were not so pure as his own.

I have also perused the Biographical Fragments of the Sieur Bode, a famous Illuminee; these will be very useful in our Historical Volume. As to numberless other works which I have read on the same subject, it will suffice to give the titles of them when quoted. I have said more than enough to show that I am not in the dark with respect to the subject on which I am writing.

I could wish to express my gratitude to those virtuous men who, by their correspondence, and the memorials which they have sent me, have greatly advanced my undertaking. But open expressions of such a gratitude would prove fatal to them. To have contributed to the public utility is a sufficient reward for their virtue; and if my work is not so perfect as it ought to be, it arises not from any want of energy in their endeavours.

I find myself much against my will obliged to answer certain objections which my Translator has made, and which will, doubtless, be repeated by many other readers, grounded on the work of Mr. Robison, entitled Proof of a Conspiracy against all the Religions and Governments of Europe, \&c. \&c. That work was published just as this Third Volume was going to the press. Its author had not then met with my two first Volumes; but in a second Edition he is pleased to mention them in his Appendix. I am much flattered by his approbation, heartily congratulate him on the zeal he has himself shown in combating the public enemy, and am happy to see that he has wrought on the best materials. Without knowing it, we have fought for the same cause with

p. 397

the same arms, and pursued the same course; but the public are on the eve of seeing our respective quotations, and will observe a remarkable difference between them. I fear lest we should be put in competition with each other, and the cause of truth suffer in the conflict. I entreat the reader to observe, that these differences arise from the different methods followed by him and myself. Mr. Robison has adopted the easiest, though the most hazardous method. He combines together in one paragraph what his memory may have compiled from many, and sometimes makes use of the expressions of the German author when he thinks it necessary. Beside, he has seen much, and read much, and relates it all together in the paragraphs marked by inverted Commas. The warning he has given in his preface will not suffice to remove the objections of some readers. In some passages he has even adopted as truth certain assertions which the correspondence of the Illuminees evidently demonstrate to have been invented by them against their adversaries, and which in my Historical Volume I shall be obliged to treat in an opposite sense. Nor will I pretend to say, that Illuminism drew its origin from Masonry; for it is a fact demonstrated beyond all doubt, that the founder of Illuminism only became a Mason in 1777, and that two years later than that he was wholly unacquainted with the mysteries of Masonry.\footnote{Original Writings, Vol. I. Let. 6, to Ajax.ï¿½Ibid. Let. 36 to M. C. Porcinsï¿½and the first Pages of the Critical History of the Degrees.}

I know perfectly well, that this will not make Illuminism less disastrous; nevertheless I am obliged to differ from Mr. Robison when treating on that subject, as well as on some other articles.ï¿½So much for objections; here is my reply.

In the first place Mr. Robison and I always agree as to the essential facts and the Conspiracy of the Illuminized Lodges; we also agree on their maxims and degrees; and this must be sufficient to convince the reader.

In the next place, in his general view of the Sect he has observed its detestable and most dangerous principles. Like a traveller he has seen the

Monstrum horrendum, informe, ingens . . .

But he has not described its forms, its manners, and its habits. Nor would it be very prudent to reject his narrative because some few circumstances are not perfectly authenticated, or because here and there some want of order may be observable.

In short, if we except one or two letters, which may be said to be translations, all the other quotations (though in the form of letters) cannot be called so, for they are not to be found in the letters of the Illuminees. They are Extracts from different parts, all brought together under one head; Mr. Robison has given them to the public in his own stile, and sometimes makes the Illuminees speak in clearer terms than is done in the Originals. His addition in the Translation of the famous letter from Spartacus to Marius, page 165-6,\footnote{See Page 4, of this Volume.} has given rise to numberless questions, how theï¿½even dï¿½was expressed in the German text. A parenthesis follows (can this mean death?). I was obliged to answer that the even dï¿½, as well as the parenthesis, were additions; but at the same time that they were not additions contrary to the

p. 398

sense of the letter. I could willingly have attributed these deviations to a difference in the editions of the Original Writings; but a new work must be supposed, as well as new letters, to justify the quotations, and all Germany must have noticed such changes. In the first place, the Court of Bavaria would have protested against such a supposition; as the Original Writings could not have coincided with an edition so dissimilar; next, the Illuminees who have not spoken in such clear language, though clear enough in their letters; in fine, the authors who have combated Illuminism, and whose quotations all exactly agree with the Edition of Munich. The Pages may change in different Editions; but whole Letters and Discourses cannot, especially when the public may, as we have seen above, have access to the Originals.

As for myself, whose name cannot be expected to have such authority as Mr. Robison's, I have taken all the precautions of which I felt myself to stand in need.\footnote{I am also afraid that the difference that exists between the degrees of Rosicrucian, of which Mr. Robison is in possession, and those which I have mentioned, may give rise to argument. I answer, 1st. That I am acquainted with three degrees of Rosicrucians, very different in themselves; 2dly. That the Cathechisms, Questions, and Rituals for the same degree greatly differ in different countries: 3dly. That I have followed the works of Mr. Lï¿½Abbe Le Franc, which Mr. Robison has quoted: 4thly. That Mr. Robison allows the degree of Knights of the Sun as described by me to be similar to that which he is in possession of. Since the publication of my Second Volume, I have received an account of the same degree which coincides with what I had said, and this degree is a sufficient ground for all that Mr. Robison or myself have asserted on the attack carried on by Masonry against Religion and Governments.}  I never make a quotation but with the Original before me; and when I translate any passage which may stagger the reader, I subjoin the original, that each may explain and verify the text. I follow the same line of conduct when I compare the different testimonies. I never mention a single law in the code without having the original before me, or the practice of it to vouch for my assertion. Hence it will be perceived, that we are not to be put in competition with each other; Mr. Robison taking a general view while I have attempted to descend into particulars: as to the substance we agree. I heartily congratulate him on his zeal in combating the monster; and though we do not agree in certain particularities, we both evince the monstrous nature of the Sect, and the certainty of its horrible Conspiracies.

p. 399

\chapter{Spartacus-Weishaupt, Founder of the Illuminees}

There sometimes appear men formed with such unhappy dispositions, that we are led to consider them in no other view than as emanations from the evil genius, bereft by the avenging God of the power of doing good. Imbecil in the sphere of wisdom, such men are only efficient in the arts of vice and destruction; they are ingenious in those conceptions, skilful in that cunning, and fruitful in those resources which enable them despotically to reign in the schools of falsehood, depravity, and wickedness. In competition with the Sophisters, these men will surpass them in the arts of exhibiting error in false and delusive colours; of disguising the vicious passions under the mask of virtue; and of clothing impiety in the garb of Philosophy. In the den of conspirators they are pre-eminent by the atrocity of their deeds; they excel in the arts of preparing revolutions, and of combining the downfal of the Altar with that of Empires. If their career be ever impeded, it is only when they approach the paths of virtue and of real science. When Heaven in its wrath permits a being of this species to appear on the earth, it has only to put nations within the sphere of his activity, and it will be awfully avenged.

With such qualities, and under such auspices, was born in Bavaria, about the year 1748, Adam Weishaupt, better known in the annals of the sect by the name of Spartacus. To the eternal shame of his Serene protector, this impious man, heretofore Professor of Law at the University of Ingolstadt, but now banished from his country as a traitor to his Prince and to the whole universe, peacefully at the court of Ernest Lewis, Duke of Saxe Gotha, enjoys an asylum, receives a pension from the public treasury, and is dignified with the title of Honorary Counsellor to that Prince.

An odious phenomenon in nature, an Atheist void of remorse, a profound hypocrite, destitute of those superior talents which lead to the vindication of truth, he is possessed of all that energy and ardour in vice which generates conspirators for impiety and anarchy. Shunning, like the ill-boding owl, the genial rays of the sun, he wraps around him the mantle of darkness; and history shall record of him, as of the evil spirit, only the black deeds which he planned or executed. Of mean birth, his youth was passed in obscurity, and but a single trait of his private life has pierced the cloud in

p. 401

which he had enveloped himselfï¿½but it is one of hateful depravity and of the most consummate villany.ï¿½Incestuous Sophister! it was the widow of his brother whom he seduced.ï¿½Atrocious father! it was for the murder of his offspring that he solicited poison and the dagger.ï¿½Execrable hypocrite! he implored, he conjured both art and friendship to destroy the innocent victim, the child whose birth must betray the morals of his father. The scandal from which he shrinks is not that of his crime; it is (he says and writes it himself) the scandal which, publishing of the depravity of his heart, would deprive him of that authority by which, under the cloak of virtue, he plunged youth into vice and error.ï¿½Monstrous Sophister! he accuses the devils of not having skreened him from this scandal by those abominations which called the vengeance of the God of Nature on the son of Judah.ï¿½Then, impudently daring, he perjures himself; he calls every thing that is sacred to witness, that neither he nor his friends ever knew of the existence of those poisons or secret means of skreening him from infamy, much less that they had ever proposed, sought, or employed them. He challenges, and at length forces, the magistrates to prove the accusation; they produce the letters of the perjured Sophister, and therein we behold him entreating a first, a second, and even a third confidant, to seek, or cause to be sought, and to communicate to him, these horrid arts. We see him recalling promises of three years standing with respect to these means. He complains of the little success of his attempts, he accuses the agents of timidity or of ignorance; he entreats and conjures them to renew their attempts, telling them, that it was not yet too late, but that expedition was necessary. Who can paint the depravity of this single trait. How monstrous the being who could have combined such depravity! That the God who humiliates the Sophister should have permitted this single trait to have been brought to light, will suffice to show how far wickedness may be carried by the man who, with virtue on his tongue, and under the shade of that sacred name, was forming and fanaticising the blood-thirsty legions of a Robespierre.

After so shocking an accusation the reader will naturally expect us to produce incontrovertible proofs. We will, therefore, first lay before him the letter of Weishaupt to his adept Hertel; it is the Third Letter in the Second Volume of the Original Writings of the Illuminees in Bavaria.

"Now," says Weishaupt to this adept, "let me, under the most profound secrecy, lay open the situation of my heart. It destroys my rest, it render me incapable of every thing. I am almost desperate. My honour is in danger, and I am on the eve of losing that reputation which gave me so great an authority over our people. My sister-in-law is with child. I have sent her to Athens (Munich) to Euriphon, to solicit a marriage licence from Rome. You see how necessary it is that she should succeed, and that without loss of time; every moment is precious. But should she fail, what shall I do?ï¿½How shall I restore the honour of a person who is the victim of a crime that is wholly mine? We have already made several attempts to destroy the child; she was determined to undergo all; but Euriphon is too timid. Yet I scarcely see any other expedient. Could I depend on Celse's secrecy (the professor Bader at Munich), he could be of great service

p. 402

to me; he had promised me his aid three years ago. Mention it to him if you think proper. See what can be done. I should be sorry that Cato knew any thing of it, lest he should tell all his friends. If you could extricate me from this unfortunate step, you would restore me to life, to honour, to rest, and to authority (that is over his people). If you cannot, I forewarn you of it, I will hazard a desperate blow, for I neither can nor will lose my honour. I know not what devil". . . . [Here decency obliges us to be silent; but he continues] "As yet nobody knows any thing of it but Euriphon; it is not too late to make an attempt, for she is only in her fourth month, and the worst of it is, that it is a criminal case, and that alone makes the greatest efforts and the most extreme (or boldest) resolution necesssary. Be well and live happier than I do, and do think of some means which can extricate me from this affair. I am yours, \&c. Spartacus."

Notwithstanding his repugnance to let Cato into the secret, Weishaupt is at length obliged to write to him on the subject, and, after repeating that which through decency we have omitted above, this monster of hypocrisy says, "what vexes me the most in all this, is that my authority over our people will be greatly diminishedï¿½that I have exposed a weak side, of which they will not fail to advantage themselves whenever I may preach morality, and exhort them to virtue and modesty."\footnote{Original Writings, Vol. I. Let. 61, to Cato.}

Now let us observe the same Weishaupt barefacedly saying in his apology, "I think and declare before God (and I wish this writing to be looked upon as a most solemn declaration), that in all my life I have never heard of those secret means (of abortion) nor of those poisons; that I have never seen nor had knowledge of any occasion when I or my friends could even have thought of advising, administering, or making any use whatever of them. And this I say in testimony and affirmation of the truth."\footnote{Introduction to his Apology, p. 6.} It is thus that by the most abominable hypocrisy he sustains a barefaced and detestable perjury.

So much for the moral virtue of this man; but our chief object is, to consider him in his character of a Conspirator. Let us then descend into that baleful abyss, and observe him in the schools of impiety, rebellion, and anarchy. Here again he appears to have been ignorant of the gradations of crime, of the space that lies between the slightest deviation from rectitude and the most profound wickedness. Here, scarcely have the magistrates cast their eyes upon him when they find him at the head of a conspiracy which, when compared with those of the clubs of Voltaire and Dï¿½Alembert, or with the secret committees of Dï¿½Orleans, make these latter appear like the faint imitations of puerility, and show the Sophister and the Brigand as mere novices in the arts of revolution. It is not known, and it would be difficult to discover, whether Weishaupt ever had a master, or whether he is himself the great original of those monstrous doctrines on which he founded his school. There exists, however, a tradition which on the authority of some of his adepts we shall lay before the reader.

According to this tradition, a Jutland merchant, who had lived some time in Egypt, began in the year 1771 to overrun Europe, pretending to initiate

p. 403

adepts in the antient mysteries of Memphis. But from more exact information I have learned that he stopped for some time at Malta, where the only mysteries which he taught were the disorganizing tenets of the antient Illuminees, of the adopted slave; and these he sedulously infused into the minds of the people. These principles began to expand, and the island was already threatened with revolutionary confusion, when the Knights very wisely obliged our modern Illuminee to seek his safety in flight. The famous Count (or rather mountebank) Cagliostro is said to have been a disciple of his, as well as some other adepts famous for their Illuminism in the county of Avignon and at Lyons. In his peregrinations, it is said, he met with Weishaupt, and initiated him in his mysteries. If impiety and secrecy could entitle a person to such an initiation, never had any man better claims than Weishaupt. More artful and wicked than Cagliostro, he knew how to direct them among his disciples to very different ends.

Whatever may have been the fact with respect to this first master, it is very certain that Weishaupt needed none. In an age when every kind of error had taken root, he did what is naturally to be expected from men who, guided by their unhappy bias, both in religious and political opinions, always select the most abominable. He must have had some notion of the ancient Illuminï¿½es, for he adopted their name, and the disorganizing principles of their horrid system. These notions were then strengthened, without doubt, by his favorite application to the disorganizing mysteries of Manichï¿½ism, since we may observe him recommending the study of them to his disciples as a preparatory step for, and as having a close connection with, those for which he was preparing them.\footnote{See the degree of Directing Illumine, oder Scottischer Ritter (Scotch Knight) page 72.}  But perfect Atheist as he was, and scorning every idea of a God, he soon despised the twofold God of Antient Illuminism, and adopted the doctrines of Manes only in as much as they threatened every government, and led to universal anarchy. He was acquainted with the systems of the modern Sophisters; but, notwithstanding all their democracy, he did not think they had given sufficient latitude to their systems of Liberty and Equality. He only adopted their hatred for God, or pure Atheism. One class led to the destruction of all civil and political laws, the other to the overthrow of all religion; he combined them both, and formed a monstrous digest, whose object was the most absolute, the most ardent, the most frantic vow to overthrow, without exception, every religion, every government, and all property whatsoever. He pleased himself with the idea of a distant possibility that he might infuse the same wish throughout the world; he even assured himself of success.

With the talents of a vulgar Sophister such a hope would have been the summit of folly; but with a genius like that of Weishaupt, formed for great crimes, it was the confidence of unlimited wickedness. The Bavarian Sophister knew his powers; he believed no crime impossible; he only sought to combine them all to reduce his systems to practice. The mediocrity of his fortune had obliged him to consecrate the latter years of his education to the study of the laws. Whether by dissimulation he concealed the plans fostered in his breast,

p. 404

or whether he had not as yet digested them all, he however found means of getting himself named to the chair of Laws in the University of Ingolstadt, before he had attained his twenty-eighth year. On the 10th of March, 1778, he writes to Zwack that he was not yet thirty years of age; and in the same letter he informs him, under secrecy, of his future projects on Illuminism, which he had founded two years before.

He must have known himself possessed of profound dissimulation; he must have been master of strange resources, to ground his plans for the subversion of all laws throughout all empires, on the very function of public interpreter of the law. It was nevertheless at the college of Ingolstadt that Weishaupt, affecting the greatest zeal for his duty, conceived himself to be admirably situated for forming and conducting by invisible means the great revolution which he had planned. He justly estimated the influence which his office of teacher gave him over his scholars, and he had the courage to supply in private the deficiency of those lessons which he was obliged to give to them in public.

But it would have been too poor a conquest for Anarchy or Impiety to have gained only those who were under the eye of the founder. Weishaupt beheld mankind subject to religious and political laws from pole to pole, and his jealous zeal weighed the means which the saints had employed to extend the faith of Christ. There still existed the scattered remnants of an Order which the imprudent policy of Kings had obliged the Sovereign Pontiff to sacrifice to the machinations of a Philosophism, the professed enemy of both Kings and Pontiffs. Weishaupt knew how to appreciate the support which the laws had acquired from men who were heretofore spread throughout all Catholic countries, and who, in the towns and villages, publicly taught youth, thundered from the pulpit against vice, directed Christians toward the path of virtue, and went to preach the faith of Christ to idolatrous and barbarous nations. He well knew how much empires were indebted to religious Orders, that in preaching the duty which each man owed to his God, strengthened the ties that bound him to his neighbour and to his Prince. Though he in his heart detested the children of Benedict, Francis, or Ignatius, he admired the institutions of these holy founders, and was particularly charmed with those of Ignatius, whose laws directed so many zealous men dispersed throughout the world toward the same object and under one head: he conceived that the same forms might be adopted, though to operate in a sense diametrically opposite.\footnote{Mirabeau de la Monarchie Prussienne, vol. V. P. 97.}  "What these men have done for the Altar and the Throne (said he to himself) why would not I do in opposition to the Altar and the Throne? With legions of adepts subject to my laws, and by the lure of mysteries, why may not I destroy under the cover of darkness, what they edified in broad day? What Christ even did for God and for Cï¿½sar, why shall not I do against God and Cï¿½sar, by means of adepts now become my apostles?"

In attributing such a wicked emulation to Weishaupt, I will not leave the historian to fruitless conjectures. No, these very wishes in plain language are contained in his confidential letters to his disciples; and he even reproaches

p. 405

them with not imitating the submission of the followers of those holy founders.\footnote{Vid. Original Writings, Vol. I. let 27, to Cato.} His most celebrated adepts have declared, that they had observed him copying them throughout his code\footnote{See the Original Writings, Vol. I Instructio pro recipientibus, art. Bï¿½Let. 2, to Ajax.ï¿½Divers letters to Cato.ï¿½Last Observation of Philo.}; they must also have remarked, that Weishaupt, in planning his systems according to the forms adopted by those religious founders, had reserved it to himself to add all the artifices which the most infernal policy could suggest. At the actual period when this conspirator formed his plans, he was ignorant of the object of Freemasonry:\footnote{See hereafter the chapter on Masonry illuminized, Vol. IV.} He only knew that the fraternity held secret meetings: he observed that they were bound by mysterious ties, and recognized each other for brethren by certain signs and words, whatever might be their country or religion. In his mind, therefore, he combined the plan of a society, which was at once to partake as much as convenient of the government of the Jesuits, and of the mysterious silence and secret conduct of Masonry. Its object was, the propagation of the most Antisocial Systems of ancient Illuminism, and of the most Antireligious Systems of modern Philosophism.

Brooding over this disastrous project, Weishaupt cast his eyes on the young pupils whom government had entrusted to his care to form them for magistrates of their country, and defenders of the laws, and he resolved to begin his warfare against both by the perversion of these youths. He beheld in distant succession his first disciples seducing others, those again, subject to his laws, forming further adepts; and thus by degrees he came complacently to view his legions multiplying and spreading from the towns to the country, and resident even in the courts of Princes. He already heard those oaths which, under the secrecy of the Lodges, were to bind the minds and hearts of those new legions who, replete with his disorganizing spirit, were silently to undermine the Altar and the Throne. He calculated the time necesssary, and smiled to think that he would one day have only to give the signal for the general explosion.

Scarcely had this modern Eratostratus attained his eight-and-twentieth year, ere he had laid the foundations of those laws which he meant to give to his disorganizing Sect. Though he had not actually written his code, he had arranged it in his mind, and he made his first essay on two of his pupils, one named Massenhausen (whom he surnamed Ajax), about twenty years of age, and afterwards a Counsellor at Burkhausen; the other called Merz (whom he surnamed Tiberius)\footnote{Weishaupt, in a letter to Zwack, says, "My three first colleagues were Ajax, you, and Merz." (Let. 15, Feb. 1778). This clearly states, that Merz was the Tiberius who was illuminated with Ajax; for it is clear that Zwack was only initiated ten months after the two adepts Ajax and Tiberius. (See Orig. Writ. Vol. I. Sect. IV.)} nearly of the same age, but whose morals and character proved so abominable, that they made even his vile seducer blush. These two disciples soon vying with their master in impiety, he judged them worthy of being admitted to his mysteries, and conferred on them the highest degree that he had as yet invented. He called them Areopagites, installed himself their chief, and called this monstrous association The Order of Illuminees.\footnote{Orig. Writ. Vol. I. Sect. IV.ï¿½Let. 2, to Philip Strozzi.}

It was on the first of May, 1776, that the inauguration was celebrated. Let the reader well observe this epoch. It indicates a feeble beginning; it preceded the French Revolution but by a few years; that however was the time when that abominable Sect first started into existence, which was to combine all the errors, all the conspiracies, and all the crimes of the adepts of Impiety,

p. 406

[paragraph continues] Rebellion, or Anarchy, and which, under the name of Jacobin, was to consummate the dreadful Revolution. Such was the origin of that Sect which I had in view when I proclaimed to all nations, and unfortunately with too much truth, "That whatever their government or religion might be, to whatever rank they might belong in civil society, if Jacobinism triumphed all would be overthrown; that should the plans and wishes of the Jacobins be accomplished, their religion with its Pontiffs, their government with its laws, their magistrates and their property, all would be swept away in the common mass of ruin! Their riches and their fields, their houses and their cottages, their very wives and children would be torn from them. You have looked upon Jacobinical faction as exhausting itself in France, when it was only making a sportive essay of its strength."\footnote{Vol. I, page 23.}

According to the wishes and intentions of this terrible and formidable Sect, nations, astonished, have yet only seen the first part of the plans formed for that general Revolution which is to beat down every Throne overturn every Altarï¿½destroy all propertyï¿½blot out every law and conclude by the total dissolution of all society!

The omen is fatal; but (more fatal still!) I have numberless proofs to demonstrate the truth of this assertion. With respect to the Conspiracies of Illuminism, I shall draw my proofs from their own code and their archives. I will begin with their code; it will lay open the object, the extent, the manner, the means and inconceivable depth of the Conspiracies of the Sect. This First Part will comprehend the plan of their conspiracies, the extract and analysis of the code of laws which they had constructed for attaining their ends. The Second Part will show their progress and their successes from their first origin, till that period when, powerful in Revolutionary Legions, without leaving their secret dens, they unite and confound themselves with the Jacobins, and in unison with them prosecute that war of desolation which menaces with total ruin the Altar of every Godï¿½the Throne of every Monarchï¿½The Law of every Societyï¿½and the Property of every Citizen. O! that I could in delineating what the Sect has done, what it is doing, and what it still meditates to do that I could but teach nations and the chiefs of nations what they themselves ought to do, to avert the impending danger; those, I say, who have mistaken these disasters for a sudden explosion, while they are in fact but an essay of the strength of the Sect, and the commencement of their general plan.

p. 407

\chapter{Code of the Illuminees - General System, and division of the code}

By the code of the sect of Illuminees I mean the principles and systems which it had formed to itself on Religion and Civil Society, or rather against all Religion and Civil Society whatever; I mean the government and the laws which it has adopted to realize its plans, and to guide the adepts in bringing the whole universe into its systems. This was not so much a code springing from an ardent mind, and an enthusiastic zeal for a great revolution, as the offspring of reflection on the means of rendering it infallible; for no sooner had Weishaupt conceived a plan, than he foresaw the obstacles which might thwart its success. Though he decorated the first pupils whom he had seduced with the title of his profound adepts, yet he did not dare unfold to them the vast extent of his plans. Pleased with having laid the foundation, he did not hurry the elevation of that edifice, which might have been exposed to fall for want of the proper precautions; no, he wished it to be as durable as time itself. For five whole years he meditated; and he foresaw that he should still have to pause for many a tedious day on the means of securing the success of his plans. His plodding head silently ruminated and slowly combined that code of laws or rather of cunning, of artifice, of snares and ambushees by which he was to regulate the preparation of candidates, the duties of the initiated, the functions, the rights, the conduct of the chiefs, and even his own. He watched every means of seduction, weighed and compared those means, tried them one after the other; and when he had adopted any of them would still reserve the power of changing them, in case he should happen to fall upon any that would be more disastrous.

Meanwhile his first disciples, now his apostles, gained him many partizans; he seduced many himself, and directed their conduct by letter. His advice was adapted to circumstances, and, artfully husbanding his promises, he kept the minds of his disciples perpetually in suspense as to the last mysteries. To his trusty adepts he promises systems of morality, of education, and of polity, all entirely new; and they might easily surmise that this future code would be no other than that of a morality without restraint, of a religion without a God, and of a polity without laws or any dependence whatsoever\footnote{Original Writings, Vol. I. Let. to Marius and Cato.}; though he did not dare

p. 409

entirely to throw away the mask. But his laws appeared imperfect, his snares were not sufficiently concealed; and he was convinced that time and experience alone could perfect the work on which he had so long meditated. Such are the colours, at least, in which we see him representing himself when his adepts, impatient to be initiated in the last mysteries, reproach him with the slowness of the proceedings: "It is from time and experience," says he, "that we are to learn. I daily put to the test what I made last year, and I find that my performances of this year are far superior. Give me then time to reflect on what may forward and on what may delay the execution of our plans; to weigh what may be expected of our people left to themselves or led and conducted by us.ÑRemember that what is done in haste, speedily falls to ruin. Leave me then to myself, let me act alone; and believe me, time and I are worth any other two."\footnote{Original Writings, Vol. I. Letters 3, 4, 47, 60 \&c to Marius and Cato.}

Let not the reader imagine that these meditations of Weishaupt alluded to the object of his views; that never varied; the destruction of Religion, the destruction of Society and the civil Laws, the destruction of property,Ñthat was the point at which he always aimed; and this impious man too well knew his crime, not to be alarmed; we see him writing to his confident, "You know the situation in which I stand. I must direct the whole by means of five or six persons. It is absolutely necessary that I should during my life remain unknown to the greater part of the adepts themselves.ÑI am often overwhelmed with the idea that all my meditations, all my services and toils are perhaps only twisting a rope or planting a gallows for myself; that the indiscretion or imprudence of a single individual may overturn the most beautiful edifice that ever was reared."\footnote{Original Writings, Vol. I. Lett. 11 and 25, to Cato.}

At other times wishing to appear above such fears, but still reproaching the adepts with want of caution, he says, "If our affairs already go on so ill, the whole will soon be undone: the fault will be thrown upon me, and, as author of every thing, I shall be the first sacrificed. Yet that is not what frightens me; I know how to take every thing on my own score; but if the imprudence of the Brethren is to cost me my life, let me at least not have to blush before men of reflection, nor to reproach myself with an inconsiderate and rash conduct."\footnote{Let. 22, to Cato.} Thus does every motive stimulate this famous Conspirator to transfuse into his code every precaution that could at the same time skreen him from condign punishment, and secure the success of his plots. At length, after five years meditation on his side, and numerous consultations with his trusty adepts, particularly with Philo, or the Baron Knigge, who acts a very exalted part in Illuminism, Weishaupt had regulated the mode of his mysteries, and had digested the code of his Sect, that is to say, the principles, the laws, and government adopted by the Illuminees to accomplish the grand object of their Conspiracy. Before we lead our readers through the immense labyrinth of this code, let us give a general idea of the system which stimulated its author to the formation of those laws.

The more we meditate on that part of the code which we shall lay before our readers when we come to treat of the mysteries of Illuminism, the more

p. 410

clearly we observe Weishaupt adopting the principles of Equality and of Liberty, (propagated by modern Philosophism) in order to present them in a new light, and to lead his disciples to the ultimate consequences of the most absolute Impiety and Anarchy.

The modern Sophisters, some following Voltaire, others Rousseau, had begun by saying, that all men were equal and free; and they had concluded with respect to Religion, that nobody, though speaking in the name of a God who reveals himself, had the right of prescribing rules to their faith. The authority of revelation being cast aside, they left no other basis for Religion to rest upon, than the Sophistry of a reason the perpetual prey of our passions. They had annihilated Christianity in the minds of their adepts. With respect to Governments they had also asserted, that all men were equal and free, and they had concluded that every citizen had an equal right to form the laws, or to the title of Sovereign; this consequence abandoning all authority to the capricious fluctuations of the multitude, no government could be legitimate but that founded on Chaos, or the volcanic explosions of the democratic and sovereign populace.

Weishaupt, reasoning on the same principles, believed both the Sophisters and the Democratic Populace to be too timid in drawing their inferences, and the following may be said to be the essence of all his mysteries.

"Liberty and Equality are the essential rights that man in his original and primitive perfection received from nature. Property struck the first blow at Equality; political Society, or Governments, were the first oppressors of Liberty; the supporters of Governments and property are the religious and civil laws; therefore, to reinstate man in his primitive rights of Equality and Liberty, we must begin by destroying all Religion, all civil society, and finish by the destruction of all property."

Had true Philosophy but gained admittance to these lodges of Illuminism, how clearly would she have demonstrated the absurdity of each and all of these principles, and the extravagance and wickedness of such consequences, both to the master and his adepts! She would have shewn, that the rights and laws of primitive man alone upon earth, or parent of a scanty generation, neither were nor ought to be the rights and laws of man living on an inhabited globe. She would have proved, that Nature, when she ordained that man should increase and multiply on this earth, and that he should cultivate it, clearly announced that his posterity were hereafter to live under the empire of social laws. She would have observed, that without property this earth would have remained uncultivated and uninhabited; that without religious and civil laws the same earth would have only nurtured straggling hordes of vagabonds and savages. Then would our Bavarian Illuminee have concluded, that his Equality and Liberty, far from being the essential rights of man in the state of perfection, would only be the instruments of his degradation, and assimilate him to the beasts of the earth, if they were to be incompatible with Property, Religion, and Society. But true Philosophy was an alien to his school; and Weishaupt, with his detestable genius formed for error, applauds

p. 411

the sophism, makes it the basis of his system, and the ultimate secret of his mysteries.

I am not simply to prove that such is the grand object of the Conspiracy, and of the ultimate revolution which he is preparing with all his adepts. Were that my only task, I should cite the blessings which the hierophant of Illuminism pours out on those hordes that roam without laws or society, and the curses which he vents against those men who, fixing their abodes, name chiefs and constituted states. The very menaces of the teacher unfold the whole of the Conspiracy. "Yes, princes and nations shall disappear from off the face of the earth; yes, a time shall come when man shall acknowledge no other law but the great book of nature: This revolution shall be the work of the secret societies, and that is one of our grand mysteries.\footnote{See hereafter the Discourse on the Mysteries.} This single passage of the code is sufficient to demonstrate both the object of the Conspiracy and the extent of the projects of the sect; but though the Conspiracy should be clearly proved, still that would be doing little for the public good. Instead of a terrible and formidable Sect, nations and chiefs of nations might mistake the Illuminees for a band of senseless madmen, plodding without means a chimerical Revolution; therefore little to be feared, and too despicable to deserve notice. Thus would wickedness find a cloak in its excesses; the Sect would prosecute its hellish plots more actively, more confidently, and more successfully, merely because their object was supposed impossible. Society would be dissolved; our laws, our religion, and our property, would be wrested from us, because we believed them proof against any attempt. Nations would tranquilly slumber on the brink of the precipice, and be plunged into destruction while they considered the fatal cause as the delusion of delirium, and smiled on the plots of Illuminism. And its founder foresaw this: for he says to his adepts, "Let the laughers laugh, let the scoffers scoff; he that compares the past with the present, will see that nature continues its course without the possibility of diverting it. Its progress is imperceptible to the man who is not formed to observe it; but it does not escape the attention of the Philosopher."\footnote{See hereafter the Discourse on the Mysteries.}

Society then calls upon me to develop more than the existence, or even the extent of the plots of the SectÑI say, it calls on me loudly to proclaim the dangers which threaten us; yes, the evils which threaten all society must be clearly shown. A manner of proceeding and an artful cunning big with crime, which will speedily plunge nations into those disasters which they may believe chimerical, is to be clearly ascertained. I have to unfold the whole of a system, an entire code, in which each institute, each maxim, each regulation, is a new step towards a universal revolution which shall strike society a mortal blow. I am not then about to inform each citizen that his religion, his country, his property, that every society, people, or nation, are menaced; unfortunately that would be a task too easily performed. But I am bound to say, "In this horrible plot, such are the dangers which threaten your country, and such the perils that hang over your persons." I must show extensive resources combined with consummate villainy, where you imagined that nothing existed but the delirium of modern Philosophism, destitute of means.

p. 412

Weishaupt, like yourselves, had foreseen numerous obstacles to his conspiracy; and it appears that he had even exaggerated them. That for which his most famous adepts seem to despise their countrymen, should be mentioned here as redounding to their honour. Weishaupt, surrounded by the faithful Bavarians, faithful to their God and to their country (rather speculating on the human heart from his books, than closely observing men in the common intercourse of life), was not aware of how very much Philosophism had forwarded his systems.\footnote{See the Last Observations of Philo.} The generation which had attained the age of manhood appeared too much infected with the antiquated ideas on religion and government. But, unfortunately, facts soon undeceived him; and this error only served, by deferring his hopes, to turn his mind to farther precautions and meditations, which sooner or later were to render his success infallible. He would say to himself, he would say to his trusty brethren, "According to my views, I cannot employ men as they are; I must form them; each class of my Order must be a preparatory school for the next; and all this must necessarily be the work of time."\footnote{Original Writings, Vol. I. Let. to Cato.} But to accelerate the time he cast his eyes on that class of young men, which, just entering the world, easily fall a prey to error, because at that age they are under the influence of their passions. I shall hereafter show what it was that both shortened the time, and abridged their education, in presenting him with whole legions of adepts ready formed to his mysteries. It is first necessary, however, that the reader should be acquainted with the profundity of his system; because, had the French Revolution not taken place, that system would alone have sufficed to render it certain and infallible; for, could the French Revolution be done away at the present moment, and the ancient regimen be restored, this code would furnish Illuminism with all the means of effectuating one that should be still more disastrous. Let us then study it, let us dissipate the cloud in which it is enveloped. Reader, your own interest requires that you should follow our steps; and observe all the snares that have been laid for you; see with what art its disciples are beguiled, with what precaution it chooses, calls, and disposes its adepts. Its proceedings appear indeed to be slow, but they are nevertheless sure. It seems to exhaust all its art to acquire a single proselyte, but the same allurements attract whole legions. Its springs are secret, but the reader must know their power and with what constancy they move toward and direct the common ruin. He has seen the people agitated, animated, and even misled to ferocity; but he must also be informed how those adepts were created who fanaticised the people and rendered them ferocious.

Weishaupt lays down as an invariable and infallible principle, that "the grand art of rendering any revolution whatsoever certainÑis to enlighten the people;Ñand to enlighten them is, insensibly to turn the public opinion to the adoption of those changes which are the given object of the intended revolution.

"When that object cannot be promulgated without exposing him that has conceived it to public vengeance, he must know how to propagate his opinion in secret societies.

p. 413

"When the object is an universal Revolution, all the members of these societies, aiming at the same point, and aiding each other, must find means of governing invisibly, and without any appearance of violent measures, not only the higher and more distinguished class of any particular state, but men of all stations, of all nations, and of every religionÑInsinuate the same spirit every whereÑIn silence, but with the greatest activity possible, direct the scattered inhabitants of the earth toward the same point." This is what he calls the grand problem on the polity of states, on which he grounds the force of secret societies, and on which the empire of his Illuminism was to rest.\footnote{See the Discourse on the Mysteries.}

"This empire once established by means of the union and multitude of the adepts, let force succeed to the invisible power. Tie the hands of those who resist; subdue and stifle wickedness in the germ;" that is to say, crush those whom you have not been able to convince.\footnote{Ibid.} He that teaches such doctrines is not to be looked on as a weak enemy. When Weishaupt reserved them for his mysteries, as well as the revelation of his ultimate object, he knew too well that they were only fitted for men who had long been trained to view them as the lessons of nature and of Philosophy, and should he meet with any who had anticipated them, it would only abridge their novitiate. But he needed nothing less than a whole generation. It was therefore to multiply the number of adepts, to dispose them by insensible degrees to receive his doctrines; by an invisible hand to direct their ideas, their wishes, their actions, and their combined efforts, that the code of laws which he framed for Illuminism constantly tended.

According to these laws, the sect is divided into two grand classes, and each of these again subdivided into lesser degrees proportionate to the progress of the adepts.

The first class is that of preparation. It contains four degrees, those of Novice, of Minerval, of Minor Illuminee or Illuminatus Minor, and of Major Illuminee or Illuminatus Major.

Some intermediary degrees belong to the class of Preparation, which may be called of Intrusion; such are those which the sect have borrowed from Freemasonry as a means of propagation. Of these masonic degrees the code of Illuminees admit the three first without any alteration: it adapts more particularly to the views of the sect the degree of Scotch Knight as an ultimate preparation for its mysteries, and it is stiled the degree of Directing Illuminee or Illuminatus Dirigens.

The second class is that of the Mysteries, and this is subdivided into the lesser and greater mysteries. The lesser comprehend the priesthood and administration of the sect, or the degrees of Priests and of Regents or Princes.

In the greater mysteries are comprized the two degrees of Magi or Philosopher, and of the Man King. The Elect of the latter compose the council and the degree of Areopagites.\footnote{See the Original Writings, Chap., II. Part II. page 8. and the last Observations of Philo, page 89, \&c. \&c.}

In all these classes, and in every degree, there is a part of the utmost consequence, and which is common to all the Brethren. It is that employment known in the code by the appellation of Brother Insinuator or Recruiter.\footnote{This is not a term of my invention; it really is to be found in the code. Insinuator or Anwerber (signifying Recruiter) are the two words generally made use of to express this character.} The

p. 414

whole strength of the Sect depends on this part; it is that which furnishes members to the different degrees; and Weishaupt, well knowing the importance of the task, turned all his genius toward it. Let us therefore begin by directing our attention to the discovery of it.

\chapter{First Part of the code of the Illuminees - Of the Brother Insinuator, or the Requiter}

By the appellation of Brother Insinuator, is to be understood the Illuminee whose peculiar office is to make proselytes for the Sect. Some brethren were more particularly instructed for that end; they might, indeed, be called the Apostles or Missionaries of the Order, being those whom the superiors sent to the different towns and provinces, and even into distant countries, to propagate its doctrines and to establish new Lodges. These had received, in addition to the common rules, farther instructions peculiar to the higher degrees. "These (as Weishaupt writes) may sometimes be the most imbecile, and at other times the most ingenious of the Brotherhood." From the former he can depend on a blind obedience to the rules he lays down, which are never to be deviated from; and with respect to the latter, provided they be zealous and punctual, should they even transgress any of the laws, it would not be in such a manner as to commit either their own safety or that of the Order; and they would soon make amends for their indiscretion by some new artifice. But, whatever may be the sense of the Illuminee, he is obliged once or twice in his life to act the part of Brother Insinuator, and that with a certain success, by the acquisition of two or three proselytes, under pain of perpetually remaining in the lower degrees. Some Brethren of high rank may have been dispensed from this formality; but as to the generality of them there exists a positive law on that point.\footnote{Original Writings. The Statutes reformed, Art. 18.} To stimulate the zeal of the Brethren, the Insinuator is by the laws of the code established superior over every novice that he has gained to the Order: It is expressed as follows: "Every Illuminee may form to himself a petty empire; and from his littleness, emerge to greatness and power."\footnote{Ibid.}

Such then is the first duty imposed upon every Illuminee for the propagation of the Sect; and this is the part which first claims our attention, in order that we may be able to form an idea of the immensurable powers of Weishaupt for seduction.

This part may be said to be subdivided into three. The rules laid down are, first, those which are to guide the Brother Insinuator in the choice of persons to be admitted or excluded; then follow those which are to teach him how to entice into the order those persons whom he has judged proper for it;

p. 416

and lastly come those rules and arts by which novices are to be formed, and even involved in Illuminism before they are officially admitted.

In order to judge of the qualifications of the persons whom he may enlist, every Illuminee is to begin by procuring tablets, which he is to keep in the form of a Journal; and this is his Diary. Assiduously prying into every thing that surrounds him, he must vigilantly observe all persons with whom he becomes acquainted, or whom he meets in company, without exception of relations, friends, enemies, or entire strangers; he must endeavour to discover their strong and their weak side; their passions and prejudices; their intimacies, and above all, their actions, interests, and fortune; in a word, every thing relating to them: and the remarks of every day he must enter in his Diary.

A twofold advantage is to be reaped from these particulars of information; first, by the Order in general and its superiors; secondly, by the adept himself. Twice every month he will make a general statement of his observations, and he will transmit it to his superiors. By these means the Order will be informed what men, in every town or village, are friendly or inimical to it. The means of gaining over the one or destroying the other will naturally occur. With respect to the Insinuator, he will learn how to judge of those who are proper persons to be received or rejected, and he will carefully insert his reasons for the admission or rejection of those persons in his monthly statements.\footnote{Original Writings.ÑThe Statutes reformed, Art. 9, 13, and following.ÑInstructions for the Insinuators, Sect. XI, No. I.Ñfor the Insinuated Nos. 1, 3, 5, \&c.ÑLet. the 4th to Ajax.}

The Recruiting Brother will carefully guard against giving the most distant hint that he is an Illuminee. This law is peremptory for the Brethren, but more particularly for all the Insinuators, whose success may often essentially depend on it. It is to them that the legislator so strongly recommends all that exterior of virtue and of perfection, that care of shunning all public scandals which might deprive them of their ascendancy over the minds of those whom they seek to entice into the Order.\footnote{See Original Writings, Vol. II. Lett. 1, and 9.} The law expressly says, "Apply yourselves to the acquiring of interior and exterior perfection;" but lest they should conceive that this perfection even hinted at the mastering of their passions, and at renouncing the pleasures of the world, he adds, "Attend particularly to the art of dissembling and of disguising your actions, the better to observe those of others, and to penetrate into their inmost thoughts. "Die kunst zu erlernen sich zu verstellen, andere zu beobachten, and aus zu forschen." It is for that reason that these three great precepts are to be found in the summary of the Code: hold thy tongueÑbe perfectÑdisguise thyselfÑalmost following each other in the same page, and serving as an explanation of each other.\footnote{Original Writings, Vol. I. p. 40. Nos. 4, 6, and 8.}

Having made himself perfect master of these precepts, and particularly of the last, the Insinuator is next to turn his attention to those persons whom he may admit or ought to reject. He is not to admit into the Order either Pagans or Jews; but he is equally to reject all religious; and above all to shun the Ex-Jesuits as he would the plague. Ordens geistliche drfen nie aufgenommen werden, und die Ex-Jesuiten soll man wie die pest fliehen.\footnote{The Last Works of Spartacus and Philo.ÑInstruction for the Stationary Prefects and Superiors, Page 153, Let. the 2d.ÑAnd Original Writings, Instructio pro Recipientibus, Nos. 1, and 5.}

The cause of such exclusions is obvious. To speak of religion, and admit, without any precaution, Jews, Turks, and Pagans, would be too open a

p. 417

manifestation of what their religion was; and not to reject religious, would be exposing themselves to be betrayed by their own adepts.

Unless they gave evident signs of a sincere amendment, all indiscreet talkers were to be rejected; and also those men whose pride, or headstrong, interested, and inconstant minds denoted that it would be impossible to infuse into them that zeal so necessary for the order; all those again, whose drunken excesses might injure that reputation of virtue which the Order was to acquire; all those, in short, whose meanness and grossness of manners would render them too untractable to give hope for their ever becoming pliant and useful.\footnote{Instructio pro Recipientibus, page 94, and Weishaupt's Letters, passim.}

"Leave those brutes, those clownish and thickheaded fellows!" he exclaims in his Chapter on Exclusions; but, though he excluded these thickheaded fellows, Weishaupt was aware that there existed a good sort of being which some might call stupid, but who are not to be told so, as advantage may be taken of their stupidity. Such were, for example, a Baron DÕErt, and many others, who holding a certain rank in the world, though destitute of common sense, have at least their riches to recommend them. "These are a good sort of beings," says our illuminizing legislator; "they are necessary beings. They augment our number and fill our coffers, augent numerum et ¾rarium. Courage then! and make these gentry swallow the bait; but beware of communicating to them our secrets; For this species of adept must always be persuaded that the degree they are in is the highest.\footnote{Original Writings. See the first Letters to Ajax and Cato.}

Indeed, there is a sort of half exclusion for princes. The Code ordains that they shall seldom be admitted, and even when they are, shall scarcely ever rise beyond the degree of Scotch Knight; or, in other words, they are never to pass the threshold of the mysteries. Hereafter we shall see the Legislator finding an expedient for introducing them beyond that degree, but still without giving them any further insight into the mysteries;\footnote{See Degree of Regent, page 154, Letter N.} and being particularly careful to hide from them certain laws of the Order.\footnote{See Instructions for the Provincial, No. 16.}

I cannot take upon myself to say, whether a similar expedient had been found as an exception to the general rule which excluded women; but it is certain, that this law was, during a long time at least, only provisional; and many of the brethren sought to revoke it. Freemasonry had its female adepts, and the Illuminees wished to have theirs. The plan is written in Zwack's own hand-writing, and he was the most intimate friend and confidant of Weishaupt, in short, his incomparable man. It is couched in the following terms:

"Plan for an Order of Women.ÑThis Order shall be subdivided into two classes, each forming a separate society, and having a different secret. The first shall be composed of virtuous women; the second, of the wild, the giddy, and the voluptuous, auschveifenden.

"Both classes are to be ignorant that they are under the direction of men. The two superiors are to be persuaded that they are under a mother Lodge of the same sex, which transmits its orders; though in reality these orders are to be transmitted by men.

"The Brethren who are intrusted with this superintendance shall forward their instructions without making themselves known. They shall conduct the

p. 418

first, by promoting the reading of good books, but shall form the latter to the arts of secretly gratifying their passions, durch begngung ihrer leidenschaften im verborgenen."

A preliminary discourse prefixed to this plan points out the object and future services of these illuminized sisters. "The advantages which the real Order would reap from this female Order would be, first, the money which the sisterhood would pay at their initiation; and, secondly, a heavy tax upon their curiosity, under the supposition of secrets that are to be learned. And this association might moreover serve to gratify those brethren who had a turn for sensual pleasure."\footnote{Original Writings, Vol. I. Sect. V.}

A list and description of eighty-five young ladies of Manheim accompanied this project of Zwack, very properly surnamed the Cato of Illuminism; from among whom, in all probability, the founders of these two classes were to be chosen. Circumstances not having favoured our modern Cato's views, we observe several other adepts proposing similar plans. An assessor of the Imperial Chamber at Wetzlaar of the name of Dittfurt, known among the Illuminees by that of Minos, and who rose to the degree of Regent, and to the dignity of Provincial, seemed to dispute the honor of this invention, both with Brother Hercules and even with Cato himself: We must allow, at least, that nobody was more anxious for the execution of the project than he was. He had already submitted his ideas to the Baron Knigge, and he applies anew to Weishaupt. He even despairs of ever bringing men to the grand object of the order without the support of the female adepts. Indeed, so ardent is his zeal, that he makes an offer of his own wife and his four daughters-in-law to be the first adepts. The eldest was exactly the person for the philosophized sisterhood; she was four-and-twenty years of age, and with respect to religion her ideas were far above those of her sex; they were modelled on her father's. He had attained to the degrees of Regent and Prince of the Illuminees, and she would have been Regent and Princess. In the higher mysteries, together with Ptolemy's wife, we should have seen the one corresponding with her father, the other with her husband. These illuminized Princesses would be the only two persons of the order who should know that they were all under the direction of men. They would preside over the trials and receptions of Minervals, and would initiate those whom they judged worthy into the grand projects of the sisterhood for the reform of governments and the happiness of mankind.\footnote{Original Writings, Vol. I. Let. of Minos, p. 169.}

But, notwithstanding all the plans and zeal of the Brethren, it does not appear that the legislator ever consented to the establishment of the Sisterhood. Yet he supplied the want of such an institution by secret instructions which he gave the Regents on the means of making the influence of women over men subservient to the order, without initiating them in any of the secrets. He says, that the fair sex having the greatest part of the world at their disposition, "no study was more worthy of the adept than the art of flattery in order to gain them; that they were all more or less led by vanity, curiosity, the pleasures or the love of novelty; that it was on that side they were to be attacked, and by that they were to be rendered serviceable to the order."\footnote{See the New Works of Spartacus and Philo, and Instructions for the degree of Regent, No. 6.} He nevertheless

p. 419

continued to exclude great talkers and women from all the degrees, nor was the sixth article of his instructions for the Insinuator rescinded.

Notwithstanding all these exclusions, the legislator leaves a sufficient scope wherein the Insinuator may exercise his zeal. He recommends generally young men of all stations from eighteen to thirty; but more particularly those whose educations were not completed, either because he thought they would more easily imbibe his principles, or would be more grateful and more zealous for doctrines for which they were indebted solely to him.\footnote{Orig. Writ. Instructiones pro Recipientibus, Page 54, No. 4; and Page 55, No. 18.}

But this preference is not an exclusion for men of a certain age, provided they are not past service, and are already imbued with the principles of Illuminism.\footnote{Ibid. Vol. II. Part the 2d, Section and Degree of Regent.} This, however, chiefly regards those persons whose rank in life can give consequence and afford protection to the order. The Recruiters are particularly instructed to insinuate themselves into the good opinion of such persons, and if possible to entice them into the Order.

There is yet another species of men, who have speech as it were at command; such are attornies, counsellors, and even physicians. "Those are worth having," says Weishaupt; "but they are sometimes real devils, so difficult are they to be led; they however are worth having when they can be gained over.\footnote{Orig. Writ. Vol. I. Let. to Ajax.}

The Insinuator is also to admit artists, mechanics of all professions, painters, engravers, white-smiths and black-smiths; but above all booksellers, those who keep post-horses, and school-masters. Hereafter the reader will see the use for which these men were intended.\footnote{See Instructions for the Insinuator, No. 4.ÑWeishaupt's Letters, passim-and the Degree of Regent.}

To yet another class of men our legislator often calls the attention of the Insinuator: "Seek me out, for example," says Weishaupt, "the dexterous and dashing youths. We must have adepts who are insinuating, intriguing, full of resource, bold and enterprising; they must also be flexible and tractable, obedient, docile, and sociable. Seek out also those who are distinguished by their power, nobility, riches, or learning, nobiles, potentes, divites, doctos, qu¾riteÑSpare no pains, spare nothing in the acquisition of such adepts. If heaven refuse its aidance, conjure hell.

"Flectere si nequeas superos, Acheronta moveto."\footnote{Ibid. Let. 3d to Ajax.}

With respect to religions, he prefers the disciples of Luther and Calvin to the Roman Catholics, and greatly prefers the former to the latter. This distinction should alone suffice to open the eyes of many who wish to persuade themselves that the whole of the revolutionary fury is aimed at the Roman Catholic religion. This motley crew certainly did the Catholics the honour of directing their shafts more pointedly at them, as strenuous opponents of their impiety and of their religious and civil anarchy; but was it to preserve the Protestant religion that Weishaupt gives them such a preference, in hopes of making them subservient to his plots? That he did give such a preference cannot be doubted, when we see him expressly writing to an adept whom he had commissioned to look out for a person proper to be received into the higher mysteries and to found a new colony of IllumineesÑwere this man a Protestant I should like him much better.ÑWre es ein Protestant,

p. 420

so wre es mir um so Lieber.\footnote{Orig. Writ. Vol. I. Let. to Tiberius, P. 223.} Weishaupt's most famous adept constantly manifests the same predilection; he even wishes to retrench certain parts of the mysteries that he may not alarm the Catholics, and seems always to hint at Frederic the IIdÕs saying, We Protestants go on brisker.\footnote{See Vol. I. page 33.} Most certainly this proves beyond a possibility of doubt, that the destruction of all Protestant laws, whether civil or religious, had place in their plans. Nor were the Protestants of Germany the dupes of such a policy, as many of the most determined antagonists of Illuminism were of that religion.

Further, he wishes to entice men into his order who have fixed residences in towns, such as merchants and canons, who might assiduously propagate his doctrines, and establish them in their neighbourhoods.\footnote{Instructions for the Provincial and Orig. Writ. Vol. I. Part II, No. 3, page 26.}

The Recruiter must use every art (for an obvious reason) to engage schoolmasters, and to insinuate his doctrines into, and gain adepts in the military academies, and other places of education; he is even to attempt the seduction of the superiors of ecclesiastical seminaries.\footnote{Ibid. Nos. 11 and 13.}

"He will spare no trouble to gain the Prince's officers, whether presiding over provinces, or attending him in his councils. He that has succeeded in this has done more," says the code, "than if he had engaged the Prince himself.\footnote{Ibid. No. 15.} In fine, the Provincial, or the chief Insinuator, is to recruit every thing that can be tainted with Illuminism, or can be serviceable to its cause."\footnote{Ibid. No. 18.}

The following extraordinary instructions are also given by Weishaupt respecting the choice of adepts: "Above all things (he says to his Insinuators) pay attention to the figure, and select the well-made men and handsome young fellows. They are generally of engaging manners and nice feelings. When properly formed, they are the best adapted for negotiations; for first appearances prepossess in their favour. It is true, they have not the depth that men of more gloomy countenances often have. They are not the persons to be entrusted with a revolt, or the care of stirring up the people; but it is for that very reason that we must know how to chuse our agents. I am particularly fond of those men whose very soul is painted in their eyes, whose foreheads are high, and whose countenances are open. Above all, examine well the eyes, for they are the very mirrors of the heart and soul. Observe the look, the gait, the voice. Every external appearance leads us to distinguish those who are fit for our school."\footnote{Let. 11th to Marius and Cato.}

"Select those in particular who have met with misfortunes, not from accidents, but by some act of injustice; that is to say, in other words, the discontented; for such are the men to be called into the bosom of Illuminism, as into their proper asylum."\footnote{Instruction for the Local Superiors, Letter H.}

Let not the reader already exclaim, How deep are the views of this illuminizing Sophister? How has he foreseen every point! With what discernment does he lay his snares to entrap those who are to be the future agents of his plots! The reader has as yet seen merely a schedule of those persons who may be admitted or rejected; but that does not sufficiently secure the order with respect to the elections which the Insinuator may have made. Before he undertakes the initiation of any person whom he may have thought

p. 421

proper, he is to make a statement from his diary of every thing that he may have observed with respect to his morals, opinions, conduct, and even of his connections in life. He is to submit this statement to his superiors, who will compare it with the notes they are already in possession of, or may acquire from other adepts, respecting the candidate, or even with a new statement, in case they judge the last to be insufficient. Even when the choice made by the Insinuator is approved of, all is not settled; the superiors have to determine which of the Insinuators is to be entrusted with the care of enticing the approved person into the Order: for all this is foreseen in the code. It is not allowed to all the brethren to exercise promiscuously so important a trust among the prophane, though they may have pointed out the person proper for reception. The young adept is not to measure his strength with the man who has the advantage over him in years and experience, nor is the tradesman to undertake the magistrate. The superior is to name the most proper Insinuator, judging from the circumstances, age, merits, dignities, or talents of the future candidate.\footnote{Instructiones pro Recipientibus. Orig. Writ. Vol. I. Nos. 1 and 7, page 54.} At length, when the mission is given, the Insinuator begins to lay his snares.ÑSuch is the second part of this extraordinary functionary, and all his subsequent steps are regulated by the code.

Candidate, in the ordinary acceptation of the word, means a person who has shown a desire or taken some steps into some order, or to acquire some dignity. In Illuminism it means the person on whom the Order has fixed its attention. It often happens, that the candidate is ignorant of the very existence of the sect. It is the Insinuator's business to inspire him with the wish of entering it. To accomplish this grand object, two different methods are inculcated. The first is for the Insinuator who has some candidate in view remarkable for his science or of a certain age. The second, for him who is entrusted with young men from eighteen to thirty, and who are susceptible of a second education. A third method was proposed for workmen, and those clownish fellows whose education had been but little attended to. We may observe Weishaupt consulting with his confidant Zwack on this part of the code; but whether it was never digested, or that he saw the Insinuators could easily supply the defect, no further mention is made of the third method. Let us then examine the essence of the first two.

To exemplify the first method, let us suppose one of those men who have gone through a complete course of modern Philosophism, who, should they not scoff at Christianity, would at least hesitate at every thing which is called religion; for the code forewarns the Insinuator, that his efforts would be vain should he attempt to seduce Philosophers of another stamp, men of sound judgment, and who would never be partizans of doctrines which could not endure the light of broad day. But when he shall have discovered one of the former who has already pretty well imbibed the principles of the sect, he will assume the character of a Philosopher well versed in the mysteries of antiquity. He will have little difficulty in acting such a part, as he will find ample instructions in the code. To follow those instructions faithfully, he must begin "by descanting on the supreme felicity of being versed in sciences which few

p. 422

can approach, of walking in the paths of light while the vulgar are groping in darkness. He must remark, that there exist doctrines solely transmitted by secret traditions, because they are above the comprehension of common minds. In proof of his assertions he will cite the Gymnosophists in the Indies, the Priests of Isis in Egypt, and those of Eleusis and the Pythagorean school in Greece."

He will select certain sentences from Cicero, Seneca, Aristides, and Isocrates; and, lest he should ever be taken unawares, he will learn those by heart which the legislator has carefully inserted in the code. Though it would be very easy to demonstrate from those very authors, that the ancient mysteries laid down no fixed principles on the important points of the Providence of God, and of the origin and order of the universe, the Insinuator is nevertheless to quote those texts to prove that there exists a secret doctrine on these objects, and above all a doctrine calculated to render life more agreeable, and pain more supportable; and to enlarge our ideas on the majesty of God. "Let him add, that all the sages of antiquity were acquainted with these doctrines; let him insist on the uncertainty that man is in with respect to the nature of the soul, its immortality, and its future destiny. He will then sound his candidate, to know whether he would not rejoice at having some satisfactory answers on objects of such great importance. At the same time he will hint that he has had the happiness of being initiated into these doctrines, and that, should the candidate wish it, he would do his best to procure him the same felicity; but that it was a science gradually imparted, and that certain men possessed the talent of guiding him from a distance, of leading him to the discovery of this new world, and that without being ever in his presence.\footnote{Original Writings, Vol. II. Part II. Sect. I.}

When the Insinuator has by such language succeeded in exciting the curiosity of his candidate, he must then ascertain his opinions on some particular articles. He will propose the discussion of certain questions in writing, and of certain principles, as the groundwork on which they are in future to proceed. The code does not determine what these questions are to be, because they vary according to the political and religious dispositions which the Insinuator may have observed in the candidate. Should these dissertations noway agree with the principles of the sect, the Insinuator will abandon his prey. Should the sophisticated candidate, or the man of importance, be found properly disposed, he will be admitted to the very threshold of the mysteries. The Insinuator will simply explain the inferior degrees to him, and mention the divers trials which the order has dispensed with in consideration of his merit.\footnote{Original Writings, Vol. II. Part II. Sect. I.}

Notwithstanding the artifice observable in this method, it is still reserved for those who need only to be acquainted with Illuminism to adopt its tenets. But should the Insinuator be entrusted with a young candidate, or with one whose principles noway coincide with those of the sect, and who is yet to be formed; it is then that Weishaupt developer that immense theory of art and cunning by which he is insensibly to ensnare his victims. "Let your first care," he says to the Insinuators, "be to gain the affection, the confidence, and the

p. 423

esteem of those persons whom you are to entice into the OrderÑlet your whole conduct be such, that they shall surmise something more in you than you wish to showÑhint that you belong to some secret and powerful society excite little by little, and not at once, a wish in your candidate to belong to a similar societyÑCertain arguments and certain books which the Insinuator must have, will greatly contribute to raise such a wish; such are, for example, those which treat of the union and strength of associations." The Legislator then carefully adds a list of those books, and the Order charges itself with the care of furnishing a certain number of them to the adepts. The works of Meiners, and particularly of Bassadows, are frequently recommended by Weishaupt, as the best fitted to inspire their readers with the love and principles of secret societies. But nothing can equal the art with which he himself has drawn up the reasons, by the help of which the Insinuator is to persuade his young candidate of the pretended necessity for these mysterious associations.

"One represents, for example," says the code, "a child in the cradle; one speaks of its cries, its tears, its weaknessÑOne remarks how this child, abandoned to itself, is entirely helpless; but that by the help of others it acquires strengthÑOne shows how the greatness of Princes is derived from the union of their subjectsÑOne exalts the advantages of the state of society over the state of natureÑThen one touches on the art of knowing and directing mankindÑHow easily, you will say, could one man of parts lead hundreds, even thousands, if he but knew his own advantages. This is evidently proved by the organization of armies, and the amazing power which princes derive from the union of their subjects."

After having descanted on the advantages of society in general, touch upon the defects of civil society, and say how little relief is to be obtained even from one's best friends,Ñand how very necessary it would be to support each other in these days. Add, that men would triumph even over heaven were they but unitedÑThat it is their disunion which subjects them to the yoke.ÑThis is to be explained by the fable of the wolf and the two dogs, the latter of whom could only be vanquished by the former after he had parted them; and by many other examples of the same kind which the Insinuator will collect.\footnote{Extract of the Instructions for the Brethren charged to enroll and receive the candidatesÑOriginal Writings, Vol. I. Sec. IX. and XII.ÑAlso in the degree of Illuminatus Major; instructions on the same object, Document A.}

As a proof of what great and important things secret societies can effectuate, he will adduce the examples of the Freemasons, of the mysterious societies of antiquity, and even of the Jesuits. He will assert, that all the great events of this world are dependent on hidden causes, which these secret societies powerfully influence; he will awake in the breast of his pupil the desire of secretly reigning, of preparing in his closet a new constitution for the world, and of governing those who think they govern us.\footnote{Original Writings, ibid. No 11 and 12. Illuminatus Major, Document A, and Letters K, L.}

"When you shall have got thus far," says the code, "begin to show (as it were unguardedly) that you are not entirely ignorant of those secrets; throw out some half sentences which may denote it. Should your candidate take the hint, press him, and return to the charge, until you see him betray symptoms of a desire instantaneously to unite with such a society.

p. 424

"The Insinuator, however, who has thus far succeeded in inspiring his pupil with such a wish, has not played off every engine with which the code has furnished him. To sound the very bottom of his mind, he will pretend to consult him as if he had been entrusted with certain secrets, he will make objections on the secrecy of these societies; but should they make too much impression he will resolve them himself. At other times, to stimulate the curiosity of his pupil, he will hold a letter in his hand written in cypher, or he will leave it half open on his table, giving his candidate sufficient time to observe the cypher, and then shut it up with all the air of a man who has important correspondences to keep secret. At other times studying the connections and actions of his pupil, he will tell him of certain circumstances which the young man will think he has learned by means of these secret societies, from whom nothing is hidden, though they are concealed from all the rest of the world."\footnote{Original Writings, ibid. No. 17 to 22.}

These artifices may be greatly abridged, according as the friendship or communicative disposition of the candidate shall have laid him more open; but on the other hand, should they not suffice, the Insinuator is not on that account to abandon his purpose; let him try to accomplish by others what he has failed in himselfÑLet him examine his own conduct, and see if he has not neglected some one or more of the rules prescribed in the code; let him redouble his attention and his complaisance. Should it be necesssary to humble himself in order to command, let not the Insinuator forget the formal precept of his legislator, "Learn also to act the valet in order to become master." Auch zu weilen den knecht gemacht, um dereinst herr zu werden.\footnote{Ibid. Let 3d, to Ajax.}

After such a long series of condescensions and discussions the candidate at length must pronounce. If he submit to all these insinuations, he is admitted among the novices of the order; but should he persist in his refusal, let him learn the fate which awaits him from those who have experienced it. "Unhappy, supremely wretched is the youth whom the Illuminees have sought in vain to entice into their sect. Should he even escape their snares, do not let him flatter himself with being proof against their hatred; and let him take care. The vengeance of secret societies is not a common vengeance; it is the hidden fire of wrath. It is irreconcilable, and scarcely ever does it cease the pursuit of its victims until it has seen them immolated."\footnote{Important Advice, \&c. by Hoffmann. Preface to Vol. II.} Such at least is the account which history gives us of those who have been guarded enough to withstand the insinuations of the Sect, and particularly of those who, after having gone the first steps with the Insinuator, have refused to proceed any farther with him.

I could cite divers examples; though I once thought that I had met with one of a quite opposite nature, in the person of Camille de Jourdan, the same deputy who was to have been involved in the sentence of transportation against Barthelemy and Pichegru after the revolution of the 4th of September, but who luckily escaped from the grasp of the triumvirate. I hear him speaking in the highest terms of one of these Insinuators, who had for a long time endeavoured to entice him into the order. He was much astonished at hearing me speak of these men as consummate in all the artifices of the most villanous

p. 425

hypocrisy. He maintained that his Illuminee was mild, modest, and moderate; full of respect for the Gospel, in a word, one of the most virtuous men he had ever known. In reply, I enumerated all the proceedings of the Insinuator, and the artifices he had played off before he quitted his prey. To all that Mr. Camille answered, "It is true; such was his behaviour; but it was his zeal for the sect which blinded him, and made him have recourse to such expedients in order to work what he called my conversion; yet, with all that, it was impossible for any body to speak of virtue and religion in so impressive a manner as he did without being at least an honest man."Ñ"Well," said I, "I will venture to assert, that the last attempt of your Insinuator was as follows. He proposed to you to give your thoughts in writing on certain questions; you did so; your opinions proved directly opposite to his; he never saw you after, became your implacable enemy, and has never since ceased calumniating you." "All that again," answered Mr. Camille, "is very true; nor was it his fault that I did not lose both friends and fortune. Before that affair he used to praise me; afterwards, however, he represented me as a most dangerous man. You cannot conceive what lies he invented about me, and I was unfortunate enough to observe that they had made impression."ÑIs it possible to be believed? Mr. Camille could not yet be persuaded but that his Insinuator was a virtuous man; so profound are the arts of hypocrisy which are to be imbibed from Weishaupt's laws! I was acquainted with two bishops, who had as completely mistaken the characters of their Insinuators as Mr. Camille de Jourdan.

But I will cite the example of Mr. Stark. I never could conceive what this Mr. Stark was whom I saw perpetually abused by the Illuminees. Nicolai and Mirabeau spared no pains to render him odious to the Protestants in Germany; they said he had received the Catholic orders of priesthood privately,\footnote{Mirabeau Monarchie Prussienne, Vol. V. art. Religion.} though every thing seemed to denote that he was a Protestant. I took some pains to inform myself who this Mr. Stark was, and I found him one of the most learned Protestant ministers in Germany; that his zeal for his religion had acquired him the degree of Doctor, and had preferred him to be Grand Almoner and Counsellor to the Landgrave of Hesse Darmstadt; but that in common with several other learned men, such as Hoffman and Zimmerman, he had had the misfortune of being sought after by the Illuminees; that he would not hearken to them; that the Illuminees had expressed a wish to have an Adept near the person of the prince, and that he had been bold enough to answer his Insinuator, "If you seek support, I am too little and my prince too great to protect you."ÑAnd every candidate who will make the same resolute stand against the agents of the order must expect to be repaid with similar calumnies. The law of the order is invariable and precise, particularly with respect to those whose talents may be obnoxious to Illuminism. They must be gained over, or ruined in the public opinion. Such is the text, so soll man den schrift steller zu gewinnen suchen, oder verschreyen.\footnote{Instructions for the Regent, No. 15.} But it is now time to follow the candidate who has shown himself more docile through the various preparatory degrees.

p. 426

\chapter{Second Part of the Code of the Illuminees - First preparatory Degree, of the Novice and of his Teacher}

In the early stages of Illuminism the duration of the time of trial for the Novice was three years for those who were not eighteen years of age; two years for those between eighteen and twenty-four; and one year for those who were near thirty.\footnote{The Statutes reformed, No. 7.} Circumstances have since occasionally caused the time to be abridged; but, whatever may be the dispositions of the Novice, though the time may be dispensed with, he must go through the different trials, or have got the start of them before he is admitted into the other degrees. During the interval he has no other superior but the Insinuator to whom he is indebted for his vocation, and during the whole time of the noviciate, the Insinuator is expressly forbidden to inform his pupil of any other member of the Order. This law was made to skreen the order from the dangers which might result from an indiscretion of the Novice, and to render the Insinuator alone responsible in such cases; for, should the Novice unfortunately be an indiscreet talker, the code expressly says, his imprudence would at most betray only one of the brethren.\footnote{The Statutes reformed, No. 16.} The first lessons of the Insinuator (in future his teacher) treat entirely on the importance and the inviolability of the secrecy which is to be observed in Illuminism. He will begin by telling his Novice, "Silence and secresy are the very soul of the Order, and you will carefully observe this silence as well with those whom you may have only reason to suppose are already initiated, as with those whom you may hereafter know really to belong to the Order. You will remember, that it is a constant principle among us, that ingenuousness is only a virtue with respect to our superiors, but that distrust and reserve are the fundamental principles. You will never reveal to any person, at present or hereafter, the slightest circumstance relative to your admission into the order, the degree you have received, nor the time when admitted; in a word, you will never speak of any object relating to the order even before Brethren, without the strongest necessity."\footnote{Original Writings, Statutes, No. 20. Statutes reformed, No. 27. True Illuminism, General Statutes, No. 31, 32.}

Under the restrictions of this severe law, one Illuminee will often be a stranger to another; and the Novice will see in this no more than a measure of safety for the order, which might be ruined by the least indiscretion.\footnote{Summary of the Statutes, No. 15. B.}

More certainly to assure himself of the discretion of the Novice, the Insinuator will give him no further insight, nor entrust him with any writing

p. 428

relative to the order, until he has obtained the following declaration: "I, the undersigned, promise upon my honour, and without any reservation, never to reveal either by words, signs, or actions, or in any possible manner, to any person whatever, either relations, allies, or most intimate friends, any thing that shall be entrusted to me by my Introducer relative to my entrance into a secret society; and this whether my reception shall take place or not. I subject myself the more willingly to this secresy, as my Introducer assures me that nothing is ever transacted in this society hurtful to religion, morals, or the state. With respect to all writings which I may be entrusted with, any letters which I may receive concerning the same object, I engage myself to return them, after having made for my sole use the necessary extracts."\footnote{Original Writings, and the true Ilium. Art. Reverse.}

These writings or books relative to the order are only lent to the Novice at first in small numbers, and for a short time; and then he must promise to keep them out of the reach of the prophane; but as he is promoted in rank, he may preserve them for a longer time, and is intrusted with a larger quantity; though not without having informed the Order of the precautions he shall have taken, lest in case of his death any of these writings should fall into prophane hands.\footnote{Institutes of the Insinuated, No 8. Orig. Writ. the real Illuminee No. 7.} He will afterwards learn, that the Brotherhood take many other precautions for secresy, not only respecting the statutes, but even with regard to the very existence of the Order. He will see, for example, in its laws, that should any of the brotherhood fall sick, the other brethren are assiduously to visit him, in the first place to fortify him, that is to say, to hinder him from making any declarations at the hour of his death; and secondly, to carry away whatever writings relative to the Order the sick man may have had in his possession, as soon as any symptoms of danger appear.\footnote{Statutes of the Minerval, No. 12.}

He will at length learn, that to frustrate all attempts to trace even their very existence, the Order does not exist every where under the same name, but that they are to assume the name of some other Order, perhaps even of a literary society, or meet without any name which can attract the attention of the public.

The first writing delivered to the Novice, to accustom him to profound secrecy, is what may be called the Dictionary of Illuminism. He must begin by learning the language of the Sect, that is to say, the art of communicating with the superiors and other adepts without the possibility of being understood by the prophane. By means of this language, the Illuminees are to be able to correspond with each other, without running the risk of its being discovered of what Brother they speak; from what place, in what language, at what period, and to whom, or by whom the letter is written.

To avoid the discovery of persons, the Novice will learn, that no Brother bears the same name in the Order which he does in the world; indeed, had he been initiated in the higher degrees of Masonry, he would have seen the same precaution taken, where the Rosicrucians receive what they call their Characteristic or their adoptive name. The Novice will receive the characteristic immediately on his admission, and it will in some measure imply the parts which he is in future to act in the general conspiracy. It will be his task

p. 429

hereafter to study and write the history of his new patron he will by this method recognize in the qualities and actions of his hero the particular services which the order will expect from him.\footnote{Original Writings, Vol. I. Instructions for the Insinuated, No. 7, and Vol. II. Let 13.} This name will be chosen as conformably as possible to the dispositions observed in him. Has he shown any propensity to repeat the impieties of Philosophism against the Gospel, he will be classed with the Celsi and Porphirii, or with the Tindals and Shaftsburys; should his turn be toward the hatred of Kings, or should his talents be judged useful for the polity of the Order, then his characteristic will be of the Brutus, Cato, or Machiavel tribe. He will not be told what he is to do to deserve his name, but they will contrive that it shall occur to him. Neither will he be told why Weishaupt assumed the name of Spartacus (a name so famous in Rome because he waged the war of the slaves against their masters); but should he ever be admitted to the higher mysteries, he will easily recognize the reason.\footnote{Original Writings, Vol. I. Sec. 4.}

The place from whence they write, as well as the persons of whom or to whom they write, is in like manner to be kept secret; a new Geography is therefore taught the Novice. He will thence learn, that Bavaria, the country of their founder, is denominated Achaia; Swabia, Pannonia; Franconia, Austria, and Tyrol are denoted by Illyria, Egypt, and Peloponnesus; Munich is called Athens; Bamberg, Antioch; Inspruck, Samos; Vienna in Austria, Rome; Wurtzburg, Carthage; Frankfort on the Mein becomes Thebes; and Heidelberg, Utica. Ingolstadt, the natal soil of the Order, was not sufficiently denoted by Ephesus; this privileged town was to be decorated with a more mysterious name, and the profound adepts bestowed on it that of Eleusis.

Should the Novice ever be sent on a mission out of his own country, or to distant shores, he will then receive further instructions in the Geography of the Sect.\footnote{Original Writings, Sect. 2 and 3.}

He must also learn how to date his letters, and be conversant with the Illuminized Hegira or Calendar; for all letters which he will receive in future will be dated according to the Persian era, called Jezdegert and beginning A.D. 630. The year begins with the Illuminees on the first of Pharavardin, which answer to the 21st of March. Their first month has no less than forty-one days; the following months, instead of being called May, June, July, August, September, and October, are Adarpahascht, Chardad, Thirmeh, Merdedmeh, Shaharimeh, Meharmeh: November and December are Abenmeh, Adameh: January and February, Dimeh, and Benmeh: The month of March only has twenty days, and is called Asphandar.\footnote{See the real Illuminee first degree.}

The Novice must next learn how to decypher the letters he may receive; in order to which, he must make himself master of that cypher, which is to serve him until initiated into the higher degrees, when he will be entrusted with the hieroglyphics of the Order.\footnote{The common cypher of the Illuminees consist in numbers corresponding to letters in the following order:

12.	11.	10.	9.	8.	7	.6.	5.	4.	3.	2.	1.

a.	b.	c.	d.	e.	f.	g.	h.	i.	k.	l.	m.

13.	14.	15.	16.	17.	18.	19.	20.	21.	22.	23.	14.

n.	o.	p.	q.	r.	s.	t.	u.	w.	x.	y.	z.

cryptogram\_image

The hieroglyphics are contained in the opposite Plate, and are copied from those published at the end of the degree of Scotch Knight or Directing Illuminee. There is a third cypher, but that has never been published.}

He will also remember, that he is never to write the name of his Order; so venerable a word cannot be exposed to prophane eyes, and a circle with a point in the middle of it will supply this sacred word, and a long square or parallelogram will denote the word Lodge.

p. 430

After these preliminary studies, the young brother receives a part of the code, under the title of Statutes of the Illuminees. But these first statutes are nothing more than a snare, and the young Novice, with pleasure no doubt, sees them begin with the following words:

"For the tranquillity and security of all the Brethren, whether Novices or active Members of the Society, and to prevent all ill-grounded suspicions, or disagreeable doubts, the venerable Order declares, that it absolutely has in view no project, enterprize, or undertaking hurtful to the state, to religion, or to good morals; and that it favours nothing of that nature in any of its members. Its designs, all its toils, solely tend to inspire men with a zeal for the perfection of their moral characters, to impregnate them with humane and sociable sentiments, to counteract the plans of the wicked, to succour oppressed and suffering virtue, to favour the advancement of men of merit, and to render those sciences universal which are as yet hidden from the generality of men. Such is not the coloured pretext, but the real object of the order."\footnote{The True Illuminee, General StatutesÑOriginal Writings, Vol. 1. Sect. 8.}

Even should the Novice not have entirely laid aside all suspicions respecting the intentions of the Order, still so positive a declaration he must think would guarantee him as to all obligations which might be imposed upon him. His grand aim is to be, to form his heart in such a maner as to gain not only the affection of his friends but even of his enemies. He is positively ordered to endeavour with all his might to acquire both interior and exterior perfection. It is true, he is soon after as positively ordered to study the arts of dissimulation and disguise; but then the Brother Insinuator is at his elbow to explain to him how that art coincides with true perfection, and thus suppress any suspicions which might arise from a comparison of these two injunctions. Beside, the Novice has many other duties to fulfil, which will deprive him of opportunity for such reflections.

He is next told, that the Brethren must have but one mind, one will, and similar sentiments; that, to effectuate this, the Order has made choice of certain works, to which he must apply with the greatest attention. Should the Novice be one of those men whom an attachment to the Gospel rendered more circumspect as to the snares laid for his belief, the very choice of the books would suffice to show him, that the first object of the Insinuator was to persuade him, that it is not even necessary to be a Christian to acquire the perfection enjoined by the statutes. The Morality he is taught is that of Epictetus, Seneca, Antoninus, and Plutarch, all foreign to Christianity. He will also receive the works of modern Sophisters, such as Wieland, Meiners, and Bassadows, who by no means make perfection to consist in Christianity. Under the soothing and mellifluous language of a moderate and specious Philosophy, he will be led to lubricity and impiety, traced by the sophisticated pen of Helvetius in his celebrated work De LÕEsprit.\footnote{See the list of these works in the Original Writings in the Statutes reformed, No. 25.} But the Insinuator must previously have sufficiently studied the dispositions of his pupil to know whether such propositions would any longer startle him. Beside, nothing is better calculated to dissipate all such fears, than the constant application that is required to those books which are put into the hands of the Novice, added

p. 431

to the care taken to deprive him of all such as might inspire him with contrary ideas. The Teacher is carefully to attend to all the rules laid down in the code on this subject, and to see that his Novices fulfil the intentions of the Order in this respect. He is frequently to converse with them; he is to mark out their occupations for them; he is even to make them unexpected visits to surprize them, and thus to see in what manner they apply to the code and other writings with which the Order has entrusted them. He is to require an account of what they have read, and extracts from the different works; he will assist them by his explanations; in short, nothing is to be neglected which can secure their progress in the spirit and morals of the Order.\footnote{See Instructiones pro Insinuantibus et Recipientibus.}

An object of far greater importance next attracts the attention of the Novice; it is that which the code calls the greatest of all; it is, the knowledge of men. The teacher will represent this to his pupil as the most interesting of all sciences.\footnote{The true Illuminee. Instructions on the Art of forming Pupils, No. 12.} To make himself master of this science, the Novice receives the model of a journal in the form of tablets, and his teacher shows him how they are to be used. Provided with this journal, he is to make his observations on every body he finds himself in company with; he is to trace their characters, and account to himself every thing he has seen or heard. Lest his memory should fail him, he must always be provided with a loose paper or small tablets, on which he may at all hours note his observations, which he is afterwards carefully to digest in his journal. To be certain of the Novice's attention to this point, the Brother Teacher will examine his tablets and his journal from time to time. To render him more expert in the art of drawing the characters of the living, he will exercise the Novices on ancient authors, and on the heroes of antiquity. No study or custom is so frequently recommended as this in all the code of Illuminism. It is to be the grand study of the Novice, and the prime occupation of every degree.\footnote{See Ibid. No. 13,ÑOriginal Writings, the Statutes, reformed, No. 9, 10, 13, 14.ÑInstructiones pro Insinuantibus, No. 5, pro Recipientibus, No. 16, \&c. \&c.}

It is by his assiduity in this great art that the Novice will learn how to distinguish those whom he may hereafter judge proper to be admitted into or rejected from the Order; and it is with that view that the Preceptor perpetually presses him to propose those whom he may think fit for the Order.\footnote{Instructiones pro Recipientibus, No 13.} By this means a double object is attained; first, the propagation of the Order; and, secondly, a knowledge of its friends or enemies; the dangers it may be threatened with; and the means to be adopted, or the persons to be gained or courted, to avert the impending storm; in fine, of extending its conquests. Whether the Illuminee be a Novice, or in any other degree, he is bound by the laws of the Order to make his report in the prescribed forms at least once a month.\footnote{Instructions for the Insinuated, No. 5. C. and Original Writings, \&c.}

While the Novice is perpetually making researches of this nature, he is not aware that he is as carefully watched by his Insinuator, who on his side notes and writes down every thing that he observes either as to the failings or the progress, the strong or weak side of his pupil, and these he as regularly transmits to the superiors.\footnote{Instructions for the Insinuator, No. 3 and 4.ÑThe real Illuminee, Instructions on the Art of forming the Brethren, No. 1, 2.}

The pupil little suspects that the grand object of his Insinuator is to bind him in such a maner to Illuminism, even long before he is acquainted with its secrets, that it shall be impossible for him to break those bonds which fear and tenor

p. 432

shall have imposed upon him, should he ever wish to shrink from the horrid plots and systems which he might thereafter discover.

This profound policy of binding the Novices to Illuminism consists, first, in giving them a magnificent idea of the grandeur of the projects of the Sect, and, secondly, in a vow of blind obedience to the superiors in every thing which they judge conducive to the ends of the Order, which vow the Insinuator is to find means of extorting from his pupil.

It is here particularly that Weishaupt appears to wish to assimilate the government of his Sect to that of the religious orders, and especially to that of the Jesuits, by a total sacrifice of their own will and judgment, which he exacts of the adepts; and to the exercising of the Novices in this point, he expressly adverts in his instructions to the Insinuators.\footnote{Mirabeau, Monarchie Prussienne, Vol. V. and Essay on the Illumines, Chap. III.ÑLast Observations by Philo, page 61.} But this is precisely the place to remark on the amazing difference between the illuminized and the religious obedience. Of that immense number of religious who follow the institutes of St. Basil, St. Benedict, St. Dominic, or St. Francis, there is not one who is not thoroughly convinced that there exists a voice far more imperious than that of his superior, the voice of his conscience, of the Gospel, and of his God. There is not one of them who, should his superior command any thing contrary to the duties of a Christian, or of an honest man, would not immediately see that such a command was a release from his vow of obedience. This is frequently repeated and clearly expressed in all religious institutes, and no where more explicitly or positively than in those of the Jesuits. They are ordered to obey their superior, but in cases only where such obedience is not sinful, ubi non cerneretur peccatum.\footnote{Constitution of the Jesuits, Part. III. Chap. I. Parag. 2, Vol. I. Edition of Prague.} It is only in cases where such obedience can have no sinful tendency whatever, ubi definiri non possit aliquod peccati genus intercedere.\footnote{Ibid. Part VI. Chap. I.} And, as if this were not sufficiently expressed, we hear their founder, at the very time when he recommends obedience to his religious, expressly saying, but remember that your vow is binding only when the commands of man are not contrary to those of God, ubi Deo contraria non pr¾cipit homo.\footnote{Epist. Ignatii De Obedientia.} All those person therefore who, like Mirabeau, surmised certain coincidences, or as he calls them points of contact, between the religious institutes and the code of the Illuminees, should have begun by observing, that religious obedience is in its very essence an obligation of doing all the good which may be prescribed without the least taint of harm. It was easy for them on the contrary to demonstrate, that the obedience sought for by Weishaupt's code was a disposition to obey every order received from the superior in spite of conscience, and unheedful of the most iniquitous guilt, provided it tended to the good of the Order. "Our society (for such are the expressions of the code) exacts from its members the sacrifice of their liberty, not only with respect to all things, but absolutely with respect to every means of attaining its end. Yet the presumption on the goodness of the means prescribed is always in favour of the orders given by the superiors. They are clearer-sighted on this object; they are better acquainted with it; and it is on this very account that they are nominated superiors It is their business to lead you through the labyrinth of

p. 433

errors and darkness; and in such a case obedience is not only a duty, but an object for grateful acknowledgment."\footnote{Statutes reformed, No. 1, 4, and 25.ÑThe true Illuminee, General Statutes, No. 11, 12.}

Such is the obedience of the Illuminees; nor is there a single exception to be found in all their code. We shall see the Novice, before he terminates his trials, obliged to explain himself explicitly with respect to orders which he may receive from his superiors, and which he may think contrary to his conscience. In the first place his teacher is to intangle him, and make himself perfectly master of his most secret thoughts. Under the pretence of knowing himself better, while studying the art of knowing others, the Novice is to draw a faithful picture of himself, to unfold his interests and connections, as well as those of his family.

Here again the Insinuator furnishes him with the tablets in the requisite form, that he may give this new proof of confidence to the Order; but this will neither be the last nor the most important one for which he will be called upon.

On these tablets, the Novice is to write down his name, age, functions, country, and abode; the species of study in which he occupies himself, the books of which his library is composed, and the secret writings of which he may be in possession; his revenue, his friends, his enemies, and the reason of his enmities; in fine, his acquaintances and his protectors.

To this table he is to subjoin a second, explaining the same objects with respect to his father, his mother, and all their other children. He is to be very explicit with respect to the education they received, to their passions and prejudices, to their strong and weak sides.

We will exemplify this second table by an extract from the Original Writings, by which the reader will perceive that parents are not very much favouredÑ"The Novice, Francis Antony St . . . . . aged 22, represents his father as violent, and of soldierlike manners; his mother as a little avaricious; the weak side of both to be flattery and interest; both living after the old fashion, and with an antiquated frankness; in their devotion, headstrong, arrogant; with difficulty abandoning an ill-conceived project, and still more unforgiving to their enemies; that they nevertheless were little hated, because little feared; and hardly in the way of doing any body any harm."

While the Novice is thus occupied in revealing all his secrets, and those of his family, the Insinuator on his side is drawing up a new statement of every thing he has been able to discover during the whole time of his pupil's trial, either with respect to him or to his relations.

On comparing the two statements, should the superior approve of the admission of the Novice to the last proofs, he is then to answer the grand questions. It is by these questions that the Novice is to judge of the extent of the sacrifice he is about to make, and of the awful subjection of his whole will, conscience, and person, to Illuminism, if he wishes to gain admittance.

The Questions are twenty-four in number, and couched in the following terms:

p. 434

I. Are you still desirous of being received into the Order of the Illuminees?

II. Have you seriously reflected on the importance of the step you take, in binding yourself by engagements that are unknown to you?

III. What hopes do you entertain, or, by what reasons are you induced to enter among us?

IV. Would you still persevere in that wish, though you should find that we had no other object or advantage whatever in view but the perfection of mankind?

V. What would be your conduct should the Order be of new invention?

VI. Should you ever discover in the Order any thing wicked, or unjust to be done, what part would you take; Wenn unanstandige, ungerechte sachen vorkamen, wie er sich verhalten wurde?

VII. Can you and will you look upon the welfare of the Order as your own?

VIII. We cannot conceal from you, that Members, entering into our Order without any other motive than to acquire power, greatness, and consideration, are not those whom we prefer. In many cases one must know how to lose in order to gain. Are you aware of all this?

IX. Can you love all the Members of the Order, even such of your enemies as may be members of it?

X. Should it so happen that you should be obliged to do good to your enemies who are of the Order, to recommend them, for example, or extol them; would you be disposed to do so?

XI. Do you, moreover, grant the power of life and death to our Order or Society? On what grounds would you refuse, or recognize in it such a right; Ob er dieser geselschaft, oder order auch das jus vit¾ et necis, aus was grnden, oder nich zugestehe?

XII. Are you disposed on all occasions to give the preference to men of our Order, over all other men?

XIII. How would you wish to revenge yourself of any injustice, either great or small, which you may have received from strangers or from any one of our Brethren?

XIV. What would be your conduct should you ever repent of having joined our Order?

XV. Are you willing to share with us happiness and misfortune?

XVI. Do you renounce the idea of ever making your birth, employment, station, or power, serve to the prejudice or contempt of any one of the Brethren?

XVII. Are you, or have you any idea of becoming a Member of any other society?

XVIII. Is it from levity, or in hopes of soon being acquainted with our constitution, that you so easily make these promises?

XIX. Are you fully determined to observe our laws?

p. 435

XX. Do you subject yourself to a blind obedience without any restriction whatever? And do you know the strength of such an engagement? Ober unbedingten gehorsam angelobe, und wisse was das sey?

XXI. Is there no consideration that can deter you from entering into our Order?

XXII. Will you, in case it is required, assist in the propagation of the Order, support it by your counsels, by your money, and by all other means?

XXIII. Had you any expectation that you would have to answer any of these questions; and if so, which question was it?

XXIV. What security can you give us that you will keep these promises; and to what punishment will you subject yourself in case you should break any of them?\footnote{Original Writings. The account of the reception of two Novices, Vol. I. Sect. 17.}

In order to judge of the nature of the answers written and signed by the Novice, and confirmed by his oath, it will be sufficient to cast our eyes on the account of the reception of two Brethren, as it is contained in the archives of the Sect. To the VIth question, should you ever discover in the Order any thing wicked, or unjust to be done, what part would you take? The first of these two Novices, aged 22, and named Francis Anthony St. . . . answers, swears, and signs, "I would certainly execute those things, if so commanded by the Order, because it may be very possible that I am not capable of judging of what is just or unjust. Besides, should they be unjust under one aspect, they would cease to be so as soon as they became a means of attaining happiness, the general end."

The Novice Francis Xaverius B. . . . answers, swears, and signs, in like manner, "I would not refuse to execute those things (wicked and unjust) provided they contributed to the general good."

To the XIth question, on life and death, the first Novice answers with the same formalities, "Yes, I acknowledge this right in the Order of Illumines; and why should I refuse it to the Order, should it ever find itself necessitated to exercise it, as perhaps without such a right it might have to fear its awful ruin. The state would lose little by it, since the dead man would be replaced by so many others. Besides, I refer to my answer to question VI.;" that is to say, where he promised to execute whatever was just or unjust, provided it was with the approbation or by order of the Superiors.

The second answers, swears, and signs to the same question, "The same reason which makes me recognize the right of life and death in the governors of nations, leads me to recognize most willingly the same power in my Order, which really contributes to the happiness of mankind as much as governors of nations ought to do."

On the XXth question, on blind obedience without restriction, one answers, "Yes, without doubt, the promise is of the utmost importance; nevertheless I look upon it as the only possible means by which the Order can gain its ends." The second is less precise: "When I consider our Order as of modern invention and as little extended, I have a sort of repugnance in binding myself by so formidable a promise; because in that case I am justified in doubting whether a want of knowledge or even some domineering passion might not sometimes occasion

p. 436

things to be commanded totally opposite to the proposed object of the general welfare. But when I suppose the order to be more universally spread, I then believe, that in a society comprehending men of such different stations, from the higher to the lower, those men are best enabled to know the course of the world, and how to distinguish the means of accomplishing the laudable projects of the Order."

This doubt of the Novice as to the antiquity of the Order must have displeased Weishaupt, who spared no pains to make it appear that Illuminism was of ancient date, the better to excite the curiosity and the veneration of the pupils; being content to enjoy the glory of his invention with his profound adepts, to whom only he revealed the secret of the invention of the highest degrees and the last mysteries. But our Novice went on to say, that on the whole he rather believed the Order to be of ancient than of modern invention; and, like his fellow Novice, he "promises to be faithful to all the laws of his Order, to support it with his counsels, his fortune, and all other means; and he finishes by subjecting himself to forfeit his honour, and even his life, should he ever break his promise.\footnote{See the two accounts.}

When the Insinuator has found means of binding the Novice to the Order by such oaths, and especially when the young candidate shall have recognized without hesitation that strange and awful right which subjects the life of every citizen to the satellites of Illuminism, should any be unfortunate enough to displease its Superiors; when the Novice is blinded to such a degree as not to perceive that this pretended right, far from implying a society of sages, only denotes a band of ruffians and a federation of assassins like the emissaries of the Old Man of the Mountain; when, in short, he shall have submitted himself to this terrible power, the oath of the modern Seyde is sent to the archives of the Order. His dispositions then prove to be such as the superiors required to confer on him the second degree of the preparatory class; and the Insinuator concludes his mission by the introduction of his pupil.

At the appointed time in the dead of the night, the Novice is led to a gloomy apartment, where two men are waiting for him, and, excepting his Insinuator, these are the first two of the Sect with whom the Novice is made acquainted. The Superior or his Delegate holds a lamp in his hand half covered with a shade; his attitude is severe and imperious; and a naked sword lies near him on the table. The other man, who serves as Secretary, is prepared to draw up the act of initiation. No mortal is introduced but the Novice and his Insinuator, nor can any one else be present. A question is first asked him, whether he still perseveres in the intention of entering the Order. On his answering in the affirmative, he is sent by himself into a room perfectly dark, there to meditate again on his resolution. Recalled from thence, he is questioned again and again on his firm determination blindly to obey all the laws of the Order. The Introducer answers for the dispositions of his pupil, and in return requests the protection of the Order for him.

"Your request is just," replies the Superior to the Novice. "In the name of the most Serene Order from which I hold my powers, and in the name of

p. 437

all its Members, I promise you protection, justice, and help. Moreover, I protest to you once more, that you will find nothing among us hurtful to Religion, to Morals, or to the State;"Ñhere the Initiator takes in his hand the naked sword which lay upon the table, and, pointing it at the heart of the Novice, continues, "but should you ever be a traitor or a perjurer, assure yourself that every Brother will be called upon to arm against you. Do not flatter yourself with the possibility of escaping, or of finding a place of security.ÑWherever thou mayst be, the rage of the Brethren, shame and remorse shall follow thee, and prey upon thy very entrails."ÑHe lays down the sword.Ñ"But if you persist in the design of being admitted into our Order, take this oath:"

The oath is conceived in the following teens:

"In presence of all powerful God, and of you Plenipotentiaries of the most high and most excellent Order into which I ask admittance, I acknowledge my natural weakness, and all the insufficiency of my strength. I confess that, notwithstanding all the privileges of rank, honours, titles, or riches which I may possess in civil society, I am but a man like other men; that I may lose them all by other mortals, as they have been acquired through them; that I am in absolute want of their approbation and of their esteem; and that I must do my utmost to deserve them both. I never will employ either the power or consequence that I may possess to the prejudice of the general welfare. I will, on the contrary, resist with all my might the enemies of human nature, and of civil society." Let the reader observe these last words; let him remember them when reading of the mysteries of Illuminism; he will then be able to conceive how, by means of this oath, to maintain civil society, Weishaupt leads the adepts to the oath of eradicating even the last vestige of society. "I promise," continues the adept, "ardently to seize every opportunity of serving humanity, of improving my mind and my will, of employing all my useful accomplishments for the general good, in as much as the welfare and the statutes of the society shall require it of me.

"I vow (ich gelobe) an eternal silence, an inviolable obedience and fidelity to all my superiors and to the statutes of the Order. With respect to what may be the object of the Order I fully and absolutely renounce my own penetration and my own judgment.

"I promise to look upon the interests of the Order as my own; and as long as I shall be a Member of it, I promise to serve it with my life, my honour, and my estates. Should I ever, through imprudence, passion, or wickedness, act contrary to the laws or to the welfare of the Serene Order, I then subject myself to whatever punishment it may please to inflict upon me.

"I also promise to help the Order, to the best of my power, and according to my conscience, with my counsels and my actions, and without the least attention to my personal interest; also, to look upon all friends and enemies of the Order as my own, and to behave to them as the Order shall direct. I am equally disposed to labour with all my might and all my means at the propagation and advancement of the Order.

p. 438

"In these promises I renounce every secret reservation, and engage to fulfill them all, according to the true purport of the words, and according to the signification attached to them by the Order when it prescribed the OathÑ

"So help me God." N. N.

The oath being signed by the Novice, and enregistered in the minutes of the Order, the Initiator declares his admission, telling him at the same time that he is not to expect to know all the members, but those only who, being of the same degree, are under the same Superior.ÑFrom that moment advanced to the degree of Minerval, he is instructed in the signs of his new degree, which are much of the same nature as those of Masonry. He is then enjoined to give an exact list of all his books, particularly of those which might be precious or useful to the Order. He also receives the following questions which he is to answer in writing.

I. What should you wish to be the object of our Order?

II. What means, either primary or secondary, do you think most conducive to the attainment of that object?

III. What other things would you wish to find among us?

IV. What men do you either hope to meet, or not to meet, among us?\footnote{True Illuminee, 1st initiation, Page 51 and following. Original Writings, Vol. I. Sect. 15.}

The answers given to these questions will enable the Superiors to judge how far the young adept has imbibed the principles of the Order. But other helps are preparing for him, that he may be able to demonstrate by his answers both the progress he has made and that which he may be expected to make.

Thus admitted to the degree of Minerval, he will find himself in future a Member of the Academy of the Sect. Let us then observe well both the Scholars and their Masters; for they still belong to the class of preparation.

p. 439

\chapter{Third Part of the Code of the Illuminees - Second preparatory Degree - The Academy of Illuminism, or the Brethren of Minerva}

\end{document}